### 2022-12-12 inst/paper/_intro.qmd

<!-- intro to calculus and residues -->
Dental calculus has proven to be excellent source of a wide variety of information
about our past.

the increased availability of methods such as aDNA, paleoproteomics, and
chromatography, has expanded our ability to identify specific pathogens, dietary,
and possible medicinal compounds<!--cite-->.

<!--how are compounds incorporated in dental calculus-->
Increased interest in using oral fluid<!--more accurate terminology than saliva?--> as a means of detecting alkaloids in living
individuals, due to the non-invasive nature of the testing compared to blood
and urine sampling<!--citations-->.

Alkaloids enter the dental calculus via two routes: (1) direct
incorporation through oral consumption of alkaloid-containing plants, whether
deliberate or accidental;
and (2) passive diffusion through the
saliva as toxins are transferred from serum to saliva [@coneInterpretationOral2007],
and then into the oral cavity through the salivary glands. The relation to serum
is why there is often a close relation between oral fluid and blood concentrations
of drugs [@coneInterpretationOral2007; @sorensenDrugsCalculus2021]
The second route allows the
detection of compounds that not only are consumed orally, but also any other pathway.
We also expect to detect metabolites through the second pathway.

Many of the components involved in the formation and growth of dental calculus
originate from saliva. Therefore the transfer of salts and other
compounds between saliva and oral biofilms on the tooth surface is likely to also
allow various alkaloids of dietary and medicinal origin to become entrapped in
dental plaque, eventually mineralising and becoming dental calculus.

Previous studies have been conducted using various chromatographic methods on
archaeological dental calculus, employing both targeted
[@eerkensDentalCalculus2018] and untargeted approaches
[@buckleyDentalCalculus2014; @gismondiMultidisciplinaryApproach2020].

<!--brief description of method validation-->
In the method validation study, all compounds detected in the blood were also
detected in the dental calculus,
with additional compounds present in dental calculus that were not present
in blood, suggesting that dental calculus represents a comprehensive history of
consumption over a long time <!-- rephrase-->[@sorensenDrugsCalculus2021].

<!--summary of results-->
We were able to detect both parent compounds, such as caffeine and nicotine, as
well as their metabolites theophylline and cotinine. These indicate the consumption
of tea and coffee and smoking of tobacco, which is also confirmed by historic
documentation and identification of pipe notches in the dentition.