### 2022-12-12 inst/paper/_methods.qmd

New dental calculus samples were taken under sterile conditions in a positive
pressure laminar flow hood in a dedicated dental calculus lab at Leiden University.
The surface of the tooth was lightly brushed with a sterile, disposable toothbrush
to get rid of surface contaminants. A sterile dental curette was then used to
scrape calculus from the tooth onto weighing paper, which was transferred to 1.5 ml
Eppendorf tubes.
Where possible, previously sampled material was used to prevent repeated sampling
of individuals. The sampling methods can be found in the published material
[@ziesemer16SChallenges2015; @ziesemerGenomeCalculus2018]. 
The samples were sent to the Department of Forensic Medicine at Aarhus University
for ultra-high-performance liquid chromatography tandem mass spectrometry (UHPLC-MS/MS).

The UHPLC-MS/MS method was validated in a separate study on cadavers received
for forensic autopsy and toxicological analysis. Results from dental calculus were
validated against compounds detected in whole blood samples from the same
individuals.

Briefly, dental calculus was treated with citric acid and the dissolution extracts
were cleaned using weak and strong polymeric cation-exchange sorbents.

[[ UHPLC method description needs to be elaborated ]]
<!--rephrase and expand-->

<!-- the samples were washed in 0.5 mL MeOH for 10 seconds, and centrifuged
at 10,000 $\times$ g for 1 minute. This process was repeated. Supernatants were
mixed with 10 mL dimethyl sulphoxide (DMSO) and 50 mL SIL-IS solution in a conical
glass centrifuge tube. The solvent was evaporated under a stream of nitrogen at
40&deg;C until ca. 10 mL residue remained, which was vortexed with 50 ml 30% (v/v)
MeOH. The calculus residue was air-dried in the dark for 24 h and weighed. The
washing/drying procedure was repeated twice using 1 mL MeOH, resulting in three
wash extracts for analysis. -->
In the original method, samples were washed three times to remove residual
substances from the surface of the calculus that originated from oral fluids,
and only extract substances from the calculus. In our samples the washes served
to remove potential contaminants from the burial environment and post-excavation
handling.
<!-- 
The remaining calculus was dissolved using lysing tube beads in 800 $\mu$L of 0.5
$\small{M}$ citric acid (CA) and 50 $\mu$L SIL-IS solution. During dissolution the
suspension was kept at ambient temperature for 1 h under continuous gentle shaking,
and subsequently mixed with 800 $\mu$L MeOH and centrifuged at 10,000 $\times$ g for
5 mins.
-->
<!--
2.4.3. Clean-up by WCX
The dissolution extract was mixed with 6300 m L of water,
100 m L of a 25% ammonia solution to obtain a pH of 6–7 and 50 m L
of the SIL-IS working solution. The solution was passed through a
WCX SPE cartridge with a maximum flow rate of 0.5 mL/min
(mainly using gravitational force (GF)). The cartridge was pre-
conditioned with 1 mL of MeOH followed by 1 mL of water. The
passed extract was collected in a glass tube for clean-up by SCX.
The cartridge was washed with 1 mL of water, then full vacuum
( 70 kPa) was applied for 5 s to remove the bulk volume of water
with a minimum of air flow passing through the sorbent. The wash
water was collected in the same tube. The substances were eluted
from the column using a 2-mL volume of freshly prepared 2% FA in
MeOH. A 10- m L volume of DMSO was added to the eluate, and then
the solvent was evaporated to a residual volume of approximately
10 m L under a stream of nitrogen at 40  C. The residue was vortexed
with a 50- m L volume of 30% MeOH, and the solution was
transferred to an autosampler vial. This solution was used for
the analysis of most substances.

2.4.4. Clean-up by SCX
The sample fraction saved from WCX clean-up was mixed with
100 m L of 12 M HCl and passed through an SCX SPE cartridge that
was pre-conditioned with 1 mL of MeOH followed by 1 mL of
water. The cartridge was washed with 1 mL of water, then full
vacuum was applied for 5 s before it was eluted with 2 mL of 5%
ammonia in MeOH. The eluate was mixed with 10 m L of DMSO
and evaporated to a residual volume of approximately 10 m L
under a stream of nitrogen at 40  C. The residue was vortexed
with a 50- m L volume of 30% MeOH, and the solution was
transferred to an autosampler vial. This solution was used for the
analysis of morphine-3-glu, morphine-6-glu, pregabalin, gaba-
pentin and salicylic acid.

2.4.5. UHPLC
A 5- m L volume was injected onto the column running 99%
mobile phase A. The column flow rate was 400 m L/min, and the
column temperature was maintained at 40  2  C. The mobile
phase was changed through a linear gradient to 0% A and 100% B
over 5 min. Nine minutes after the initial injection, the gradient
was returned to 99% A over 0.1 min, and the column was
equilibrated for 2.9 min before the next injection, resulting in a
total run time of 12 min. The same column and gradient
programme were used for the wash, WCX and SCX extracts. The
eluent was diverted to waste during the time intervals of 0–1.5 and
8–12 min after injection using a post-column switch.

2.4.6. MS-MS
The probe temperature (TEM) was set to 400  C. The curtain
gas (CUR), ion source gas 1 (GS1), ion source gas 2 (GS2) and
collision gas (CAD) were set at 20, 80, 80 and 8 psi, respectively.
The mass spectrometer was operated in both positive ion mode
and negative ion mode at unit mass resolution using polarity
switching to obtain the maximum sensitivity for both analytes
that favour positive ionisation and analytes that favour negative
ionisation. The ion spray voltages were set at 4.5 kV. Scheduled
multiple reaction monitoring (MRM) with windows of 30 s was
applied, and at least 12 data points were obtained across each
peak. The applied MRM conditions are listed in Supplementary
Table S2. Nitrogen was used as the CAD gas. Data acquisition and
processing were performed using Analyst 1.7 and MultiQuant
3.0.3 (Sciex), respectively. -->

For more details about the method and validation, see the original study by
Sørensen and colleagues [-@sorensenDrugsCalculus2021].

Of the `r nrow(demography)` individuals that were analysed,
`r nrow(filter(metadata, replicated == TRUE))` individuals were subsequently
replicated using leftover calculus from the original DNA study.
in <!-- the same lab? different lab? -->.

### Skeletal analysis

<!-- methods for age, sex, and stature estimation -->
Demographic and pathological analyses were conducted in the Laboratory for Human
Osteoarchaeology at Leiden University. Sex was estimated using cranial
and pelvic morphological traits [@Standards1994]. Age was estimated using dental wear,
auricular and pubic surface appearance, cranial suture closure, and epiphyseal fusion 
[@Standards1994; @SucheyBrooks1990; @buckberryAuricular2002; @lovejoyAuricular1985; @meindlSutureClosure1985].
<!-- differential diagnoses -->

#### Paleopathology

Various pathological conditions were included in the analysis and were
dichotomised to presence/absence to allow statistical analysis.
Osteoarthritis (OA) was considered present in cases with eburnation.
No distinction was made between active or <!-- ? --> lesions in cribra femora
and cribra orbitalia.
Data on chronic maxillary sinusitis were obtained from another study
[@casnaUrbanizationRespiratory2021].

[[ Description of method for sinusitis - Maia ]]
<!-- Description of lesions included in path conditions - Supp Material?

- OA = osteoarthritis
- IVDD = intervertebral disc disease
- TB = tuberculosis
- DISH = diffuse idiopathic skeletal hyperostosis
- VOP = vertebral osteophytosis
- SN = schmorl's nodes
- DDD = degenerative disc disease
- PNBF = periosteal new bone formation
- OD = osteochondritis dissecans
- CF = cribra femora
- CO = cribra orbitalia

Sinusitis as well
-->

#### Dental pathology

Caries ratios were calculated by dividing the number of lesions by the number
of teeth scored, resulting in a single caries ratio per individual. If the surface
where the lesion originated is not visible, i.e. if the lesion covered multiple
surfaces, this was scored as "crown".

Calculus indices were calculated according to Greene and colleagues [-@greeneQuantifyingCalculus2005].
Calculus was scored with a four-stage scoring system (0-3) to score absent, slight,
moderate, and heavy calculus deposits [@brothwellDiggingBones1981]
on the lingual, buccal (labial), and interproximal surfaces of each tooth. Only
one score was used for the combined interproximal surfaces, resulting in three
scores per tooth (when surfaces are intact), and four calculus indices per individual;
upper anterior (UA), upper posterior (UP), lower anterior (LA), lower posterior (LP).
Each index was calculated by dividing
the sum of calculus scores for each surface by the total number of surfaces scored
in each quadrant. If a tooth could not be scored on all three surfaces, the tooth
was not included [@greeneQuantifyingCalculus2005].

### Statistical analysis

All compounds and pathological conditions/lesions were converted to a presence/absence
score. Pearson product-moment correlation was applied to the
dichotomised pathological lesions (point-biserial correlation),
compound concentrations, calculus indices, and caries ratios to explore relationships
paired continuous-continuous variables and paired continuous-binary variables.
Compound concentrations were then dichotomised to presence/absence, and
the caries ratio and calculus index for each individual were converted to an
ordinal score from 0 to 4 by using quartiles. Polychoric correlation was applied
to the paired dichotomous variables and dichotomous-ordinal variables.
<!--Median periodontal scores were calculated for each individual. -->

All statistical analysis was conducted in `r version$version.string`,
`r version$nickname`, [@Rbase]. Data wrangling was conducted with the **tidyverse**
[@tidyverse2019] and visualisations were created using **ggplot2** [@ggplot2].
Polychoric correlations were calculated with the **psych** package [@Rpsych].