### 2022-12-12 inst/paper/_results.qmd

<!-- detected compounds -->
Multiple compounds were detected in the dental calculus samples. Compounds detected
at a lower concentration than the LLOQ were considered not present. Not all the
compounds detected in the first batch could be replicated in the second batch
(@tbl-compound-detect). For a full list of targeted compounds, see Supplementary
Material.
<!-- maybe move full list to supp. mat and only include compounds here that were detected in one of the batches. -->

```{r}
#| label: detection-tbl-setup

uhplc_calculus_detect <- uhplc_calculus_long %>%
#uhplc_calculus_detect <- uhplc_calculus_demography %>%
  #remove_missing() %>%
  ungroup() %>% 
  filter(quant > 0) %>%
  distinct(compound, batch) %>%
  pivot_wider(names_from = batch, values_from = batch) %>% 
  mutate(across(!compound, ~ if_else(!is.na(.x), TRUE, FALSE)))
```

<!-- May need to limit the list to compounds that were detected in either batch and keep the full list in supplementary materials -->
```{r}
#| label: tbl-compound-detect
#| tbl-cap: "Target compound including whether it was detected (TRUE) or not (FALSE) in each batch, as well as the lower limit of quantitation (LLOQ) in ng."
uhplc_calculus_detect %>%
  left_join(compound_name_abbrev, by = c("compound" = "abbrev")) %>% 
  full_join(lloq, by = c("name" = "compound")) %>%
  mutate(across(batch1:batch2, ~ replace_na(.x, FALSE))) %>%
  filter(!(batch1 == FALSE & batch2 == FALSE)) %>% 
  select(name, batch1, batch2, lloq) %>% 
  arrange(name) %>% 
  knitr::kable(col.names = c("Compound", "Batch 1", "Batch 2", "LLOQ"))
```

Most plots show a large increase in extracted mass of a compound between the
calculus wash extracts (wash 1-3) and the dissolved calculus (calc). Most samples
containing theophylline and caffeine had the largest quantity of the compound
extracted from the first wash, then decreasing in washes 2 and 3. There is
an increase between wash 3 and the dissolved calculus in all samples.
The patterns are consistent across batches 1 and 2.
Nicotine and cotinine have the same relative quantities in the samples, i.e., the
sample with the highest extracted quantity of nicotine also had the highest extracted
quantity of cotinine [@fig-auth-plot-batch2].

```{r}
#| label: fig-auth-plot-batch2
#| fig-cap: "Quantity (ng) of each compound extracted from each sample across washes 1-3 and final calculus extraction (calc). Colours represent different calculus samples. Only results from batch 2 are shown."

uhplc_data_long %>%
  filter(
    batch == "batch2",
    !compound %in% c("cbd", "cbn", "cocaine", "thc", "thca-a", "thcva") # remove compounds not detected in batch 2
    ) %>% 
  semi_join(quant_filter, by = c("sample", "compound")) %>% # remove compounds not detected in each sample 
  mutate(
    extraction = factor(extraction, levels = c("wash1", "wash2", "wash3", "calc")),
    #sample = as.factor(sample)
    ) %>% 
  ggplot(aes(x = extraction, y = quant, group = sample, colour = sample)) +
    geom_line() +
    geom_point(size = 0.2) +
    #facet_wrap(~ compound, scales = "free_y", labeller = compound_name_repair) +
    facet_wrap(~ compound, scales = "free_y", labeller = labeller(compound = compound_names)) +
    theme_bw() +
    theme(
      legend.position = "none"
    ) +
    labs(y = "Quantity (ng)", x = "") +
    scale_colour_viridis_d()
```

<!-- detection vs. preservation -->
To see if preservation of the skeletal remains had any effect on the detection of
compounds, concentrations of compounds were compared to the various levels of
preservation. No apparent patterns were present between preservation and detection
of compounds in the samples (@fig-detection-preservation).

```{r}
#| label: fig-detection-preservation
uhplc_calculus_long %>%
  filter(
    !is.na(preservation),
    !compound %in% c("cbd", "cbn", "cocaine", "thc", "thca-a", "thcva"), # remove compounds not detected
    id != "MB329" # unusually high nicotine and cotinine quantity in batch 1
    ) %>% 
  ggplot(aes(x = preservation, y = quant)) +
    geom_violin(aes(fill = preservation), alpha = 0.6) +
    #geom_jitter(width = 0.2) +
    geom_boxplot(width = 0.2) +
    facet_wrap(~ compound, scales = "free_y", labeller = labeller(compound = compound_names)) +
    theme_bw() +
    theme(legend.position = "none") +
    labs(x = "Preservation", y = "Quantity (ng)") +
    scale_fill_viridis_d()
```

<!-- detection of tobacco as accuracy -->

```{r}
#| label: tobacco-accuracy-setup
#tobacco <- uhplc_calculus_demography %>%
tobacco <- uhplc_calculus_long %>%
  filter(compound %in% c("nicotine", "cotinine")) %>% 
  mutate(
    detection = if_else(quant == 0, 0, 1) # 0 = not detected; NA = not included in batch 2.

    ) # 0 = not detected; NA = not included in batch 2.

tobacco_accuracy <- tobacco %>%
  remove_missing(vars = "quant") %>%
  group_by(sample, id, .drop = F) %>% 
  summarise(
    detection = sum(detection),
    #compound = compound
    ) %>% 
  left_join(select(demography, id, pipe_notch, preservation, age), by = "id") %>% 
  mutate(
    pipe_notch = if_else(pipe_notch > 0, "Y", "N"),
    correct = case_when(
      detection == 0 & pipe_notch == "N" ~ 1,
      detection > 0 & pipe_notch == "Y" ~ 1,
      detection > 0 & pipe_notch == "N" ~ NaN, # no way of knowing macroscopically if the person smoked without pipe
      TRUE ~ 0
    )
  )

accuracy_age <- tobacco_accuracy %>% # move to supplementary material
  group_by(age) %>% 
  summarise(mean = mean(correct, na.rm = T))

# accuracy in replicated samples
replicated_accuracy <- tobacco %>%
  group_by(sample,id) %>%
  summarise(detection = sum(detection)) %>%
  remove_missing() %>% 
  left_join(select(demography, id, pipe_notch, preservation, age), by = "id") %>% 
  mutate(
    pipe_notch = if_else(pipe_notch > 0, "Y", "N"),
    correct = case_when(
      detection == 0 & pipe_notch == "N" ~ 1,
      detection > 0 & pipe_notch == "Y" ~ 1,
      detection > 0 & pipe_notch == "N" ~ NaN, # no way of knowing macroscopically if the person smoked without pipe
      TRUE ~ 0
    )
  )
```

The presence of pipe notch(es) in an individual and concurrent detection of nicotine
and/or cotinine is used as a crude indicator of the accuracy of the method. When
combining the results of both batches, the method was able to detect some form
of tobacco in
`r nrow(filter(tobacco_accuracy, pipe_notch == "Y", detection > 0))`
of `r nrow(filter(tobacco_accuracy, pipe_notch == "Y"))`
individuals with a pipe notch
(`r scales::percent(nrow(filter(tobacco_accuracy, pipe_notch == "Y", detection > 0)) / nrow(filter(tobacco_accuracy, pipe_notch == "Y")), accuracy = 0.1)`).
When also considering correct the absence of a tobacco alkaloid together with the absence
of a pipe notch, the accuracy of the method is
`r scales::percent(mean(tobacco_accuracy$correct, na.rm = T), accuracy = 0.1)`.
Accuracy in the old adult age category is
`r scales::percent(filter(accuracy_age, age == "old")$mean, accuracy = 0.1)`.

In the replicated samples only, tobacco detection was successful in
`r nrow(filter(replicated_accuracy, pipe_notch == "Y", detection > 0))`
out of
`r nrow(filter(replicated_accuracy, pipe_notch == "Y"))`
pipe smokers
(`r scales::percent(nrow(filter(replicated_accuracy, pipe_notch == "Y", detection > 0)) / nrow(filter(replicated_accuracy, pipe_notch == "Y")), accuracy = 0.1)`)
Including individuals with absence of a pipe notch and concurrent absence of
compounds as a correct identification, gives an overall accuracy of
`r scales::percent(mean(replicated_accuracy$correct, na.rm = T), accuracy = 0.1)`.

One individual---an old adult, probable female---was positive
for both nicotine and cotinine, and had no signs of a pipe notch.

<!-- missing data -->

Additional results from dental and skeletal lesions can be found in the supplementary
materials.

### Correlations between detected alkaloids and diseases

For further statistical analyses, only the UHPLC-MS/MS results from batch 2 were
used, as batch 1 had results that could not be replicated and may have contained
contamination.

```{r}
#| label: tbl-pearson
corr_tib <- conc_cor %>% 
  as_tibble(rownames = "var") %>% 
  filter(var != "age",
         var != "periodont_status") %>% 
  column_to_rownames("var") %>%
  mutate(across(everything(), signif, 3))
corr_tib %>% 
  knitr::kable(caption = "Pearson correlation on dichotomous skeletal lesions and compound concentrations from batch 2. Correlations between pairs of dichotomous variables are removed due to incompatibility with a Pearson correlation.")
```

Point-biserial (Pearson) correlation on caries and calculus versus
presence/absence variables. We found only weak correlations between the calculus
index and caries ratio and other variables.

Polychoric correlation was conducted on the dichotomised compounds and pathological
conditions, as well as the discretised dental diseases.
Salicylic acid was removed due to its ubiquitous presence in the sample, and is
likely to cause spurious correlations.
Strong correlations were found between `r cor_statement(polycorr_tib, strength = "strong")`.
Moderate correlations were found between `r cor_statement(polycorr_tib, strength = "moderate")`. Remaining
correlations were weak or absent (@fig-polycorr).
Correlations with age will be depressed because age was largely controlled for
in the sample selection.

```{r}
#| label: fig-polycorr
#| fig-cap: "Plot of the polychoric correlations. Larger circles and stronger calculus indicate stronger correlation coefficient."
polycorr$rho %>%
  ggcorrplot::ggcorrplot(
    method = "circle", type = "lower", 
    colors = c(viridisLite::plasma(1, begin = 0), "white", viridisLite::plasma(1, begin = 1))
    ) +
  scale_x_discrete(labels = compound_names) +
  scale_y_discrete(position = "right") +
  theme(
    legend.position = "left",
  )
```