### 2022-12-12 inst/paper/paper.qmd

---
title: "Title Goes Here"
author:
  - Bjørn Peare Bartholdy:
      correspondence: "yes"
      email: seriousname@email.org
      orcid: 0000-0003-1689-0557
      institute:
        - fosg
  - Jørgen B. Hasselstrøm
  - Lambert K. Sørensen
  - Maia Casna
  - Historisch Genootschap Beemster
  - Amanda G. Henry
institute:
  - fosg:
      name: ZZ9 Plural Z Alpha
      address: 23 Science Street, Eureka, Mississippi, USA
  - fop: Federation of Planets
title-block-published: "Last updated"  
#date: now
#date-format: long
format: 
  html: 
    theme: cosmo
execute: 
  freeze: true
knitr: 
  opts_chunk:
    echo: false
    warning: false
    message: false
    comment: "#>"
    fig-path: "../figures/"
    fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  #- ../templates/pagebreak.lua
bibliography: references.bib
csl: "apa.csl" # Insert path for the bib-style
abstract: |
  This is an abstract.
keywords: |
  stuff; and; things
highlights: |
  These are the highlights. 
---

```{r}
#| label: setup
#| include: false
#| cache: false

library(here)
library(readr)

# upload data
metadata <- read_tsv(here("inst/data/raw_data/metadata.tsv"))
demography <- read_csv(here("inst/data/raw_data/demography.csv"))
lloq <- read_tsv(here("inst/data/raw_data/lloq.tsv"))
uhplc_data_comb <- read_csv(here("inst/data/derived_data/uhplc-data_combined.csv"))
#uhplc_calculus <- read_csv(here("inst/data/derived_data/uhplc-calculus_cleaned.csv"))
dental_inv <- read_csv(here("inst/data/raw_data/dental-inv.csv"))
caries <- read_csv(here("inst/data/raw_data/caries.csv"))
periodont <- read_csv(here("inst/data/raw_data/periodontitis.csv"))
periap <- read_csv(here("inst/data/raw_data/periapical.csv"))
calculus <- read_csv(here("inst/data/raw_data/calculus.csv"))
calculus_full <- read_csv(here("inst/data/raw_data/calculus_full.csv"))
sinusitis_clean <- read_csv(here("inst/data/derived_data/sinusitis_cleaned.csv"))
path_cond_clean <- read_csv(here("inst/data/derived_data/path-conditions_cleaned.csv"))

source(here("inst/scripts/setup-qmd.R"))
#source(here("inst/supplementary-materials/supp-mat.qmd"), local = knitr::knit_global())
# set viridis as default pallette

options(ggplot.discrete.colour = function() scale_fill_viridis_d())
```

Keywords: {{< meta keywords >}}

<!-- Highlights: {{< meta highlights >}} -->

<!-- The actual document text starts here: -->

## Introduction

{{< include _intro.qmd >}}

## Materials

{{< include _materials.qmd >}}

## Methods

{{< include _methods.qmd >}}

## Results

{{< include _results.qmd >}}

## Discussion

{{< include _discussion.qmd >}}

## Conclusion

## Acknowledgements {-}

Cats and dogs

## References {-}

::: {#refs}
:::