### 2022-12-12 inst/supplementary-materials/supp-mat.qmd

---
title: "Supplementary Materials"
knitr: 
  opts_chunk: 
    warning: false
    message: false
format: 
  html:
    toc: true
---

```{r}
#| label: setup
#| include: false
library(here)
library(readr)
devtools::load_all()

# upload data
metadata <- read_tsv(here("inst/data/raw_data/metadata.tsv"))
demography <- read_csv(here("inst/data/raw_data/demography.csv"))
lloq <- read_tsv(here("inst/data/raw_data/lloq.tsv"))
uhplc_data_comb <- read_csv(here("inst/data/derived_data/uhplc-data_combined.csv"))
#uhplc_calculus <- read_csv(here("inst/data/derived_data/uhplc-calculus_cleaned.csv"))
dental_inv <- read_csv(here("inst/data/raw_data/dental-inv.csv"))
caries <- read_csv(here("inst/data/raw_data/caries.csv"))
periodont <- read_csv(here("inst/data/raw_data/periodontitis.csv"))
periap <- read_csv(here("inst/data/raw_data/periapical.csv"))
calculus <- read_csv(here("inst/data/raw_data/calculus.csv"))
calculus_full <- read_csv(here("inst/data/raw_data/calculus_full.csv"))
sinusitis_clean <- read_csv(here("inst/data/derived_data/sinusitis_cleaned.csv"))
path_cond_clean <- read_csv(here("inst/data/derived_data/path-conditions_cleaned.csv"))

source(here("inst/scripts/setup-qmd.R"), local = knitr::knit_global())
```

## Session information

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r}
print(sessionInfo(), locale = F)
```

## Data dictionary

### metadata

| variable | description |
|---|---|
| id | unique identifier for the individual |
| sample | sample number for the UHPLC-MS/MS analysis |

### dental-inv

Dental inventory

| variable | description |
|---|---|
| feat | Identifier: Feature number of individual |
| find | Identifier: Find number of individual |
| t11..t48 | Tooth number (FDI system) and status |

Dental inventory key:

- p = present
- m = missing (for unknown reason)
- aml = ante-mortem loss
- dna = removed for DNA sampling

### caries

Caries lesions location

| variable | description |
|---|---|
| feat | Identifier: Feature number of individual |
| find | Identifier: Find number of individual |
| t11..t48 | Tooth number (FDI system) and location of caries |

Caries location key:

- none = no caries present
- mes = mesial
- dis = distal
- lin = lingual (and labial)
- buc = buccal
- occ = occlusal
- crown = crown (caries-related loss of much of/entire tooth crown)
- root = root
- blank = tooth not present

### periodontitis

| variable | description |
|---|---|
| feat | Identifier: Feature number of individual |
| find | Identifier: Find number of individual |
| t11..t48 | Tooth number (FDI system) and periodontitis score (0-3) |

Periodontitis score:

- 0 = none
- 1 = slight
- 2 = moderate
- 3 = severe

- blank = tooth not present

### periapical

| variable | description |
|---|---|
| feat | Identifier: Feature number of individual |
| find | Identifier: Find number of individual |
| t11..t48 | Tooth number (FDI system) and periapical lesion |

- none = no lesion
- bucc = buccal
- lin = lingual
- perf = perforated alveolar bone
- blank = not scoreable

### calculus

| variable | description |
|---|---|
| feat | Identifier: Feature number of individual |
| find | Identifier: Find number of individual |
| t11..t48 | Tooth number (FDI system) and deposit size (0-3) |

- 0 = no calculus
- 1 = slight calculus
- 2 = moderate calculus
- 3 = heavy calculus

Scored according to Standards.

### path-conditions

Pathological conditions

- OA = osteoarthritis
- IVDD = intervertebral disc disease
- TB = tuberculosis
- DISH = diffuse idiopathic skeletal hyperostosis
- VOP = vertebral osteophytosis
- SN = schmorl's nodes
- DDD = degenerative disc disease
- PNBF = periosteal new bone formation
- OD = osteochondritis dissecans
- CF = cribra femora
- CO = cribra orbitalia

## Sample demographics

```{r}
#| label: male-sample
males <- demography %>%
  filter(sex == "m" | sex == "pm")
```

The sample consists of `r nrow(demography)` individuals, most of which are
middle adult male individuals (@fig-age-distribution).

```{r}
#| label: fig-age-distribution
demography %>%
  ggplot(aes(x = age, fill = sex)) +
    geom_bar() +
    theme_minimal()
```

## Missing data

Missing data per dataset

```{r}
#| label: fig-missing-hex
caries %>%
  mutate(across(!id, ~ is.na(.x))) %>%
  select(!id) %>%
  colSums()

caries %>%
  mutate(across(!id, ~ is.na(.x))) %>% 
  pivot_longer(-id) %>% 
  ggplot(aes(x = name, y = id, fill = value)) +
    geom_tile()

```

Missing data across all dental datasets

```{r}
#| label: fig-missing-dental
dental_long %>%
  mutate(across(c(caries, periodont, periap, calculus), is.na)) %>%
  select(c(caries, periodont, periap, calculus, id, tooth)) %>%
  #group_by(id, tooth) %>% 
  rowwise() %>% 
  mutate(missing = sum(caries, periodont, periap, calculus)) %>% 
  ggplot(aes(x = tooth, y = id, fill = missing)) +
    geom_tile()
```

## UHPLC analysis

```{r}
#| label: tbl-lloq
#| tbl-cap: "Target compounds and lower limits of quantitation."
lloq %>%
  arrange(compound)
```

```{r}
#| label: tbl-uhplc-batch-1
#| tbl-cap: "Results from the UHPLC analysis first batch. Quantity of compound in the dissolved calculus, represented in ng."
uhplc_calculus_long %>%
  filter(batch == "batch1") %>%
  #select(sample, contains("batch1")) %>%
  #rename_with(~ str_remove(.x, "_calc_batch1"), .cols = contains("_calc_batch1"))
  pivot_wider(names_from = "compound", values_from = "quant") %>%
  select(!c(extraction, batch))
```

```{r}
#| label: tbl-uhplc-batch-2
#| tbl-cap: "Results from the UHPLC analysis second batch. Quantity of compound in calculus after third wash, represented in ng."
uhplc_calculus_long %>%
  filter(batch == "batch2") %>%
  #select(sample, contains("batch1")) %>%
  #rename_with(~ str_remove(.x, "_calc_batch1"), .cols = contains("_calc_batch1"))
  pivot_wider(names_from = "compound", values_from = "quant") %>%
  select(!c(extraction, batch))
```

### Authentication

No modern synthetic drugs were detected in any of the samples.

Samples were replicated to verify results from the initial analysis. Of the
`r nrow(demography)` samples initially analysed,
`r nrow(filter(metadata, replicated == TRUE))` samples were replicated.

Only caffeine, theophylline, nicotine, cotinine, and salicylic acid were found
in the replicated samples.

```{r}
#| label: tbl-uhplc-replicated
uhplc_calculus_replicated %>%
  filter(sum(presence) != 0) %>% # remove compounds absent in all individuals in sample
  pivot_wider(names_from = "compound", values_from = "presence")
```

<!-- line plot with concentration across washes and to the final calculus sample -->

Most plots show a large increase in extracted mass of a compound between the
calculus wash extracts (wash 1-3) and the dissolved calculus (calc). Most samples
containing theophylline and caffeine had the largest quantity of the compound
extracted from the first wash, then decreasing in washes 2 and 3. There is
an increase between wash 3 and the dissolved calculus in all samples.
The patterns are consistent across batches 1 and 2. The pattern we expect to see
in a sample is a reduction in the quantity from wash 1 to wash 3, and then another
spike in the final extraction from the dissolved calculus, which means the compound
is actually 'ancient' or authentic. The compounds that are completely absent in
all three washes and present in high quantities in the final extraction may also
be suggestive of lab contamination.
This has not been thoroughly tested and is
only based on what we expect to see. Therefore, the interpretation of these graphs
is itself up for interpretation.

```{r auth-plot-batch1}
#| label: fig-auth-plot-batch1

uhplc_data_long %>%
  filter(
    batch == "batch1",
    ) %>% 
  semi_join(quant_filter, by = c("sample", "compound")) %>% 
  mutate(
    extraction = factor(extraction, levels = c("wash1", "wash2", "wash3", "calc")),
    sample = as.factor(sample)
    ) %>% 
  ggplot(aes(x = extraction, y = quant, group = sample, colour = sample)) +
    geom_line() +
    geom_point(size = 0.2) +
    facet_wrap(~ compound, scales = "free_y") +
    theme_bw()
```

```{r}
#| label: fig-auth-plot-batch2
uhplc_data_long %>%
  filter(
    batch == "batch2",
    !compound %in% c("cbd", "cbn", "cocaine", "thc", "thca-a", "thcva") # remove compounds not detected in batch 2
    ) %>% 
  semi_join(quant_filter, by = c("sample", "compound")) %>% # remove compounds not detected in each sample 
  mutate(
    extraction = factor(extraction, levels = c("wash1", "wash2", "wash3", "calc")),
    sample = as.factor(sample)
    ) %>% 
  ggplot(aes(x = extraction, y = quant, group = sample, colour = sample)) +
    geom_line() +
    geom_point(size = 0.2) +
    facet_wrap(~ compound, scales = "free_y") +
    theme_bw()
```

### Quantity vs. sample weight

There is no clear relationship between the sample weight and the amount of compound
detected, except for salicylic acid, where the amount of extracted compound increases
with increasing sample weight. In batch 2 there is also a slight positive trend
for caffeine, nicotine, and cotinine.
Nicotine and cotinine display the same relative relationship between samples. Where
the nicotine quantity is high compared to other samples, the cotinine quantity
will be similarly high (@fig-quant-weight-1 and @fig-quant-weight-1).

```{r}
#| label: fig-quant-weight-1
#| #| fig-cap: "Quantity of a compound (ng) found in a sample plotted against the weight of the calculus sample. Results from batch 1."
uhplc_data_comb %>%

  select(sample, contains("batch1")) %>% 
  select(sample, batch1_weight, contains("calc")) %>% 
  pivot_longer(
    -c(sample, batch1_weight),
    names_to = c("compound", "batch"),
    names_pattern = "(.*)_(.*)",
    values_to = c("conc")
  ) %>% 
  filter(conc > 0) %>% 
  mutate(compound = str_remove(compound, "_calc")) %>%
  ggplot(aes(x = batch1_weight, y = conc, col = as.factor(sample))) +
    geom_point() +
    facet_wrap(~ compound, scales = "free_y") +
    theme_bw() +
    theme(legend.position = "none") +
    labs(x = "", y = "Quantity (ng)")
```

```{r}
#| label: fig-quant-weight-2
#| fig-cap: "Quantity of a compound (ng) found in a sample plotted against the weight of the calculus sample. Results from batch 2."
uhplc_data_comb %>%
  select(sample, contains("batch2")) %>% 
  select(sample, batch2_weight, contains("calc")) %>% 
  pivot_longer(
    -c(sample, batch2_weight),
    names_to = c("compound", "batch"),
    names_pattern = "(.*)_(.*)",
    values_to = c("conc")
  ) %>% 
  mutate(compound = str_remove(compound, "_calc")) %>%
  #remove_missing() %>% 
  filter(conc > 0) %>% 
  ggplot(aes(x = batch2_weight, y = conc, col = as.factor(sample))) +
    geom_point() +
    facet_wrap(~ compound, scales = "free_y") +
    theme_bw() +
    labs(y = "Quantity (ng)", y = "", col = "Sample number")
```

### Distribution of compounds detected in the samples

<!-- absolute counts in each batch -->

```{r}
#| label: fig-compounds-detect
#| fig-cap: "Number of individuals in which each compound was detected between batch 1 and 2."

# awkward wrangling required because the information on which individuals were included in replication was lost when combining batches 1 and 2
#n_batch1 <- length(metadata$replicated)
#n_batch2 <- sum(metadata$replicated) # sum boolean vector to get length of TRUE values
uhplc_calculus_long %>%
  #remove_missing() %>%
  filter(quant > 0) %>% 
  #mutate(detect = replace_na(conc, 0)) %>% 
  #mutate(detect = if_else(detect > 0, 1, 0)) %>%
  #group_by(compound, batch) %>% 
  # summarise(
  #   n_detect = sum(detect)
  #           ) %>% 
  # mutate(n = case_when(batch == 1 ~ n_batch1,
  #                      batch == 2 ~ n_batch2)) %>% 
  # mutate(prop = n_detect / n) %>%

ggplot(aes(x = compound, fill = compound)) +
    #geom_jitter(width = 0.2) +
    #geom_boxplot() +
    geom_bar() +
    facet_wrap(~ batch) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
      legend.position = "none"
      )
```

The replication showed that caffeine, theophylline, cotinine, nicotine, and
salicylic acid could be consistently detected in the samples, although theophylline
detection decreased between batches 1 and 2. CBD, CBN, cocaine, and THCA-A was not
detected at all in the second batch.

<!-- absolute counts in the replicated individuals -->

```{r}
#| label: fig-compounds-detect2
#| fig-cap: "Number of individuals in which each compound was detected between batch 1 and 2. Only showing replicated individuals."

uhplc_calculus_long %>%
  filter(id %in% filter(metadata, replicated == T)$id,
         quant > 0) %>%
  #remove_missing() %>%
  ggplot(aes(x = compound, fill = compound)) +
    geom_bar() +
    facet_wrap(~ batch) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
      legend.position = "none"
      )
```

### Detection and preservation

To see if preservation of the skeletal remains had any effect on the detection of
compounds, absolute quantities of compounds were compared to the various levels of
preservation.

```{r}
#| label: fig-detection-preservation
uhplc_calculus_long %>%
  filter(
    !is.na(preservation),
    !compound %in% c("cbd", "cbn", "cocaine", "thc", "thca-a", "thcva") # remove compounds not detected
    ) %>% 
  ggplot(aes(x = preservation, y = quant, fill = preservation)) +
    geom_violin() +
    geom_jitter(width = 0.2) +
    #geom_boxplot(alpha = 0.6, width = 0.2) +
    facet_wrap(~ compound, scales = "free_y") +
    theme_bw()
# 
# uhplc_calculus_demography %>%
#   filter(
#     !is.na(preservation),
#     !compound %in% c("cbd", "cbn", "cocaine", "thc", "thca-a", "thcva") # remove compounds not detected
#     ) %>% 
#   ggplot(aes(x = preservation, y = quant, fill = preservation)) +
#     geom_violin() +
#     geom_jitter(width = 0.2) +
#     #geom_boxplot(alpha = 0.6, width = 0.2) +
#     facet_wrap(~ compound, scales = "free_y") +
#     theme_bw()
```

```{r}
#| label: fig-detection-preservation2
#uhplc_calculus_demography %>%
uhplc_calculus_long %>% 
  filter(!is.na(preservation)) %>% 
  ggplot(aes(x = preservation, y = quant, fill = preservation)) +
    geom_boxplot() +
    facet_wrap(~ compound, scales = "free_y") +
    theme_bw()
```

### Detection of tobacco

Given that pipe notches are present in the majority of individuals, the presence
of pipe notch(es) in an individual and concurrent detection of nicotine and/or
cotinine is used as a rough indicator of the accuracy of the method.

```{r}
#| label: fig-corr-notches
# correlation between nicotine and cotinine conc, and pipe notches
uhplc_calculus_long %>%
  ggplot(aes(x = pipe_notch, y = conc)) +
    geom_point() +
    facet_wrap(~ compound + batch, scales = "free_y")
```

```{r}
tobacco <- uhplc_calculus_long %>%
  filter(compound %in% c("nicotine", "cotinine")) %>% 
  mutate(detection = if_else(is.na(conc), FALSE, TRUE))

tobacco_comb <- tobacco %>% 
  group_by(batch, sample, id, .add = TRUE) %>%
  summarise(
    detection = sum(detection), 
    #sex, age, completeness, preservation, pipe_notch, 
    .groups = "keep" # combine compounds: 0 = none detected; 1 = 1 detected; 2 = both detected
  ) %>%
  ungroup() %>%
  left_join(select(demography, id, pipe_notch), by = "id") %>%
  mutate(
    pipe_notch = if_else(pipe_notch > 0, "Y", "N"),
    correct = case_when(
      detection > 0 & pipe_notch == "Y" ~ TRUE,
      detection == 0 & pipe_notch == "N" ~ TRUE,
      detection > 0 & pipe_notch == "N" ~ NA, # can't be sure if this is true or not
      TRUE ~ FALSE
      )
    )

nicotine <- tobacco %>%
  filter(compound == "nicotine") %>% 
  mutate(
    pipe_notch = if_else(pipe_notch > 0, "Y", "N"),
    correct = case_when(
      detection > 0 & pipe_notch == "Y" ~ TRUE,
      detection == 0 & pipe_notch == "N" ~ TRUE,
      detection > 0 & pipe_notch == "N" ~ NA, # can't be sure if this is true or not
      TRUE ~ FALSE
    )
  )

cotinine <- tobacco %>%
  filter(compound == "cotinine") %>% 
  mutate(
  correct = case_when(
    detection > 0 & pipe_notch == "Y" ~ TRUE,
    detection == 0 & pipe_notch == "N" ~ TRUE,
    detection > 0 & pipe_notch == "N" ~ NA, # can't be sure if this is true or not
    TRUE ~ FALSE
    )
  )

# accuracy in replicated samples
uhplc_accuracy <- uhplc_calculus_replicated %>%
  left_join(select(demography, id, pipe_notch), by = "id") %>% 
  filter(compound == "nicotine" | compound == "cotinine") %>%
  group_by(id) %>% 
  arrange(desc(presence)) %>%
  slice(which(compound == "nicotine" || compound == "cotinine")) %>%
  mutate(
    pipe_notch = if_else(pipe_notch > 0, "Y", "N"),
    correct = case_when(presence == 1 & pipe_notch == "Y" ~ TRUE,
                             presence == 0 & pipe_notch == "N" ~ TRUE, # can we be sure if this is correct?
                             presence == 1 & pipe_notch == "N" ~ NA, # can't be sure if this is true or not
                             TRUE ~ FALSE))

# ratio of nicotine to cotinine
tobacco_ratio <- tobacco %>%
  select(id, batch, compound, conc) %>% 
  pivot_wider(names_from = compound, values_from = conc) %>% 
  group_by(id, batch) %>% 
  summarise(ratio = cotinine / nicotine) %>%
  remove_missing() %>% 
  left_join(select(demography, id, age, sex, preservation, pipe_notch)) %>% 
  filter(batch == "batch2",
         ratio != Inf) %>% 
  mutate(
    age = case_when(
      age == "eya" ~ 0,
      age == "lya" ~ 1,
      age == "ma" ~ 2,
      age == "old" ~ 3
    ),
    preservation = case_when(
      preservation == "fair" ~ 0,
      preservation == "good" ~ 1,
      preservation == "excellent" ~ 2
    )
  )

tobacco_ratio %>%
  ggplot(aes(y = ratio, x = pipe_notch)) +
    #geom_boxplot()
    geom_point()

  #lm(ratio ~ age + sex + preservation + pipe_notch, tobacco_ratio)
```

No apparent correlation between the number of pipe notches
and the concentration of nicotine or cotinine [@fig-corr-notches].

Nicotine and cotinine detection successfully identified
`r nrow(filter(tobacco_comb, batch == 1, pipe_notch == "Y", detection > 0))`
smokers out of
`r nrow(filter(tobacco_comb, batch == 1, pipe_notch == "Y"))`
(`r scales::percent(nrow(filter(tobacco_comb, pipe_notch == "Y", batch == 1, detection > 0)) / nrow(filter(tobacco_comb, batch == 1, pipe_notch == "Y")), accuracy = 0.1)`)
individuals with pipe notches in the first analysis.

Nicotine detection successfully identified
`r nrow(filter(nicotine, batch == 1, pipe_notch == "Y", detection > 0))`
smokers out of
`r nrow(filter(nicotine, batch == 1, pipe_notch == "Y"))`
individuals with pipe notches
(`r scales::percent(nrow(filter(nicotine, batch == 1, pipe_notch == "Y", detection > 0)) / nrow(filter(nicotine, batch == 1, pipe_notch == "Y")), accuracy = 0.1)`).

Cotinine detection successfully identified
`r nrow(filter(cotinine, batch == 1, pipe_notch == "Y", detection > 0))`
smokers out of
`r nrow(filter(cotinine, batch == 1, pipe_notch == "Y"))`
individuals with pipe notches
(`r scales::percent(nrow(filter(cotinine, batch == 1, pipe_notch == "Y", detection > 0)) / nrow(filter(cotinine, batch == 1, pipe_notch == "Y")), accuracy = 0.1)`).

In the replicated samples only, tobacco detection was successful in
`r nrow(filter(uhplc_accuracy, pipe_notch == "Y", presence > 0))`
out of
`r nrow(filter(uhplc_accuracy, pipe_notch == "Y"))`
smokers
(`r scales::percent(mean(filter(uhplc_accuracy, pipe_notch == "Y")$correct, na.rm = T))`).
Including individuals with absence of a pipe notch and concurrent absence of
compounds as a correct identification, gives an overall accuracy of
`r scales::percent(mean(uhplc_accuracy$correct, na.rm = T), accuracy = 0.1)`.

One individual---an old adult, probable female---was positive
for both nicotine and cotinine, and had no signs of a pipe notch.

## Dental analysis

```{r}
#| label: fig-dental-plots
#| layout-ncol: 2
#| fig-cap: 
#|   - "Antemortem tooth loss in sample."
#|   - "Periodontal scores in the sample."
dental_long %>%
  dental_plot(fill = status)

dental_long %>%
  filter(!is.na(periodont)) %>% 
  dental_plot(fill = periodont)
```

### AMTL

### Caries

Caries were scored based on location in each individual. Multiple locations on
a single tooth is possible. The size of caries was also
recorded, but not used in further analysis. Large caries that cover multiple
surfaces were recorded as 'crown'. Caries counts were extracted based on the
number of locations recorded.

| code | surface |
|---|---|
| mes | mesial surface |
| dis | distal surface |
| lin | lingual surface |
| buc | buccal surface (including labial surface) |
| occ | occlusal surface (including incisal surface) |
| crown | caries covers 2+ surfaces |
| none | No caries visible on surface |
| NA | Not observable/tooth missing |

```{r}
surface <- c("mes", "dis", "occ", "buc", "lin", "root", "crown") # here occ = incisal

# caries_count <- dental_long %>% 
#   select(!c(periodont, periap, calculus)) %>% 
#   #filter(status == "p") %>%
#   separate_rows(caries, sep = ";") %>% # one lesion per row
#   mutate(caries_count = if_else(
#     caries == "none", 0L, 1L
#     )
#   ) %>%
#   group_by(id, tooth) %>%
#   summarise(
#     caries_count = sum(caries_count, na.rm = F)
#     ) %>%
#   tooth_position()
# 
# site_wide_caries <- caries_count %>%
#   caries_ratio(.caries = caries_count)
```

In the `r caries_ratio_site$n_teeth` that were examined, `r caries_ratio_site$count`
teeth had caries
(`r scales::percent(caries_ratio_site$ratio, accuracy = 0.1)`).
This frequency has very little meaning, and was further broken down into tooth
class and individuals.

```{r}
caries_count %>%
  group_by(class) %>% 
  caries_ratio(.caries = count)
```

As expected, the molars have a higher frequency of caries than the other teeth.

```{r}
caries_count %>% 
  group_by(id, class) %>% 
  caries_ratio(.caries = count)
```

```{r}
caries_count %>%
  caries_ratio(.caries = count, id, class)
```

```{r}
caries_count %>%
  caries_ratio(.caries = count, class, id) %>%
  ggplot(aes(x = class, y = ratio)) +
    geom_violin(aes(fill = class)) +
    geom_boxplot(width = 0.1)
```

```{r}
upper_order <- c(paste0("t", 18:11), paste0("t", 21:28))
lower_order <- c(paste0("t", 48:41), paste0("t", 31:38))

maxilla <- caries_count %>%
  filter(region == "maxilla") %>% 
  mutate(tooth = factor(tooth, levels = c(upper_order, lower_order))) %>% 
  group_by(tooth) %>% 
  summarise(
    n_teeth = n(),
    count = sum(count, na.rm = T),
    rate = count / n_teeth
    ) %>% 
  ggplot(aes(x = tooth, y = rate)) +
    geom_col(fill = "white") +
    theme_dark() +
    scale_y_reverse(limits = c(0.33,0), sec.axis = sec_axis(~.)) +
    scale_x_discrete(position = "top") +
    theme(
      axis.title.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.line = element_line(colour = "red", size = 1),
      axis.line.x.top = element_line(colour = "red", size = 4)
      ) +
  labs(y = "")

mandible <- caries_count %>%
  filter(region == "mandible") %>% 
  mutate(tooth = factor(tooth, levels = c(upper_order, lower_order))) %>% 
  group_by(tooth) %>% 
  summarise(
    n_teeth = n(),
    count = sum(count, na.rm = T),
    rate = count / n_teeth
    ) %>% 
  ggplot(aes(x = tooth, y = rate)) +
    geom_col(fill = "white") +
    scale_y_continuous(sec.axis = sec_axis(~.)) +
    theme_dark() +
    theme(axis.line = element_line(colour = "red", size = 1),
          axis.line.y.right = element_line(colour = "red", size = 1),
          axis.line.x.bottom = element_line(colour = "red", size = 6),
          axis.ticks.x = element_blank()) +
    labs(y = "Caries rate")

maxilla / mandible + plot_layout(guides = "collect")
```

### Periodontitis

```{r}
#| label: fig-periodont
dental_long %>%
  filter(!is.na(periodont)) %>% 
  dental_plot(fill = periodont)
```

### Calculus

```{r}
calc_index <- calculus_full %>%
  dental_longer(-id) %>% 
    #   group_by(id, quadrant) %>%
    # summarise(
    #   n_surf = sum(!is.na(score)),
    #   score_sum = sum(score, na.rm = T),
    #   calc_index = score_sum / n_surf
    # )

  calculus_index()
```

Distribution of calculus indices within the sample.
<!-- it is not calculating per surface (I think) -->

```{r}
#| label: fig-calculus-index
calc_index %>%
  ggplot(aes(x = quadrant, y = calc_index)) +
    geom_violin(aes(fill = quadrant), alpha = 0.6) +
    geom_boxplot(width = 0.1) +
    theme_minimal() +
    theme(panel.grid.major.x = element_blank())
```

```{r}
calc_index %>%
  left_join(uhplc_calculus_long, by = "id") %>%
  filter(
    batch == "batch1",
    compound != "thc",
    compound != "cbd",
    compound != "thcva",
    quadrant == "LA"
    ) %>% 
  ggplot(aes(x = as.factor(presence), y = calc_index)) +
    geom_violin(aes(fill = as.factor(presence)), alpha = 0.6) +
    geom_boxplot(width = 0.1) +
    facet_wrap(~ compound)
```

No apparent influence of lower anterior calculus index on the presence/absence
of a compound (or vice versa).

## Pathological conditions

## Statistical analysis

### Point-biserial correlation

Point-biserial (Pearson) correlation was conducted on compound concentrations,
calculus index, caries ratio, and binary variables.
This is done to see if any correlations exist prior to discretisation of continuous
variables.

```{r}
#| label: fig-pearson-corr
conc_cor %>%
  as_tibble(rownames = "var") %>% 
  mutate(across(!var, ~ if_else(.x == 1, NaN, .x))) %>%
  column_to_rownames("var") %>% 
ggcorrplot::ggcorrplot()
```

### Polychoric correlation

The calculus index and caries ratio for each individual was converted to an
ordinal variable by using quartiles, providing a score from 0--4.

```{r caries-calculus-quartiles}
caries_discrete
calculus_discrete
```

```{r polycorr}
polycorr$rho
```

```{r fig-polycorr}
ggcorrplot::ggcorrplot(polycorr$rho, type = "lower", show.diag = F)
```