### 2023-01-07 inst/paper/_conclusions.qmd
incorporate feedback from coauthors

This preliminary study outlines the benefits of using calculus to target a
variety of compounds that could have been consumed as medicine or diet. 
This method allows us to directly address specific individuals, which can be
especially useful in individuals that are not always well-documented in historic
documentation, such as rural communities, children and women.
We also
show that there are many limitations that will need to be addressed going forward
with this type of analysis, and stress the need for more systematic research on
the consumption of alkaloid-containing items and their subsequent concentration
and preservation in dental calculus,
in addition to how mode of consumption may affect concentrations on different parts
of the dentition.
Another limitation of dental calculus as a medium is the inter- and intra-individual
variability of its formation and the many factors that can influence incorporation
and retention of molecules and particles; however, in the absence
of hair and serum (quite uncommon in archaeology), dental calculus represents
an impressive long-term reservoir of information regarding the consumption of various
alkaloids, whether dietary, medicinal, recreational, or otherwise.