### 2023-01-07 inst/paper/_intro.qmd

<!-- intro to calculus and residues -->
Dental calculus has proven to be an excellent source of a wide variety of information
about our past. The increased accessibility and advancement of methods such as
aDNA, paleoproteomics, and chromatography, has expanded our ability to identify
biomarkers of diet and disease on an increasingly large scale
[@warinnerEvidenceMilk2014; @velskoDentalCalculus2017; @gismondiMultidisciplinaryApproach2020].

<!-- define alkaloids -->
One such biomarker is alkaloids, a plant-derived group of compounds. Many alkaloids
have important medicinal and psychoactive effects in humans, and their direct
detection, or detection of their metabolites, is of great interest to archaeologists.
Previous studies have successfully recovered alkaloids archaeological
contexts including ceramics [@smithDetectionOpium2018], pipes
[@raffertyCurrentResearch2012], human hair
[@ogaldeIdentificationPsychoactive2009; @echeverriaNicotineHair2013], and even
dental calculus employing both targeted [@eerkensDentalCalculus2018] and
untargeted approaches
[@buckleyDentalCalculus2014; @gismondiMultidisciplinaryApproach2020].
Especially nicotine, the principal alkaloid in tobacco leaves,
has been widely studied in the archaeological record due to its apparent stability
and ability to survive over long periods of time
[@eerkensDentalCalculus2018; @raffertyCurrentResearch2012; @tushinghamHuntergathererTobacco2013].

<!--how are compounds incorporated in dental calculus-->
Alkaloids may enter the oral cavity via two pathways: (1) direct
incorporation through oral consumption of alkaloid-containing plants, whether
deliberate or accidental;
and (2) passive diffusion as alkaloids and other compounds are transferred from
serum to saliva, and then into the oral cavity through the salivary glands
in the hours-to-days following consumption [@coneInterpretationOral2007]. The
relation to serum
is why there is often a close correlation between presence (not concentration)
of drugs in oral fluid and blood/serum
[@coneInterpretationOral2007; @willeRelationshipOral2009; @milmanOralFluid2011].
The second pathway allows the
detection of compounds that not only are consumed orally, but also any other mode
of consumption or administration, as long as it is excreted through saliva<!--cite e.g. injection or topical The second pathway may also be more
detectable, as the excretion of compounds through saliva may take hours to days
whereas compounds spend less time in the oral cavity during direct consumption-->.

Many of the components involved in the formation and growth of dental calculus
originate from oral fluid. Proteins, bacteria, salts and other compounds are
transferred from saliva to biofilms on the tooth surface
[@whiteDentalCalculus1997; @jinSupragingivalCalculus2002].
This may also allow various alkaloids of dietary and medicinal origin to become
incorporated in dental plaque. Dental plaque undergoes frequent mineralisation
events, ultimately causing the entrapped alkaloids and their metabolites to become
preserved within the dental calculus. Barring intentional or accidental removal of
the calculus during life,
burial, excavation, and post-excavation cleaning, the alkaloids can then be detected
by various methods to show a record of consumption during life.

<!--brief description of method validation-->
In this study we use a ultra-high-performance liquid chromatography-tandem mass
spectrometry method using pneumatically assisted electrospray ionisation
(UHPLC-ESI-MS/MS) that was developed in a previous study on dental
calculus from cadavers and validated by comparing the results to compounds detected
in the blood of the same individuals.
All compounds that were detected in the blood were also detected in dental calculus,
with additional compounds present in dental calculus that were not present
in blood, suggesting that dental calculus represents a comprehensive history of
consumption over a long period of time [@sorensenDrugsCalculus2021].
<!--summary of results-->
We were able to detect both parent compounds and metabolites, including caffeine,
nicotine, theophylline, and cotinine, in the dental calculus of individuals from
a 19th century Dutch population from Middenbeemster. By detecting these compounds
we are able to show the consumption
of tea and coffee and smoking of tobacco on an individual scale, which is also
confirmed by historic documentation and identification of pipe notches in the
dentition.