### 2023-01-29 inst/paper/_methods.qmd

New dental calculus samples were taken under sterile conditions in a positive
pressure laminar flow hood in a dedicated dental calculus lab at Leiden University.
The surface of the tooth was lightly brushed with a sterile, disposable toothbrush
to get rid of surface contaminants. A sterile dental curette was then used to
scrape calculus from the tooth onto weighing paper, which was transferred to 1.5 ml
Eppendorf tubes.
Where possible, previously sampled material was used to prevent repeated sampling
of individuals. The sampling methods can be found in the published material
[@ziesemer16SChallenges2015; @ziesemerGenomeCalculus2018]. 
The samples were sent to the Department of Forensic Medicine at Aarhus University
for ultra-high-performance liquid chromatography-tandem mass spectrometry (UHPLC-MS/MS)
analysis.

The list of targeted compounds included both naturally occurring compounds known
to have been used in the past, as well as synthetic modern drugs that did not exist
at the time (e.g. Fentanyl, MDMA, Amphetamine). These were part of the toxicology
screening for the original method [@sorensenDrugsCalculus2021],
developed on cadavers. In our study they serve as an authentication step, as their
presence in archaeological samples could only be the result of contamination.

Briefly, samples of dental calculus were washed three times each with one mL of
methanol (MeOH), to remove surface contaminants. The wash solutions were collected
separately. The solvent was
evaporated and the residues were dissolved in 50 µL 30% MeOH.
The washed calculus was homogenized in presence of 0.5 M citric acid using a lysing
tube with stainless steel beads. Following one hour of incubation the dissolution
extract was cleaned by weak and strong cation-exchange. After evaporation of the
elution solvent the residue was dissolved in 50 µL 30% MeOH.
The final extracts obtained from washing and dissolution of the dental calculus were
analysed by UHPLC-MS/MS using a reversed-phase biphenyl column for chromatography.
To obtain quantitative results, isotope dilution was applied.
For more details about the method and validation, see the original study by
Sørensen and colleagues [-@sorensenDrugsCalculus2021].

### Skeletal analysis

<!-- methods for age, sex, and stature estimation -->
Demographic and pathological analyses were conducted in the Laboratory for Human
Osteoarchaeology at Leiden University. Sex was estimated using cranial
and pelvic morphological traits [@Standards1994]. Age-at-death was estimated using
dental wear,
auricular and pubic surface appearance, cranial suture closure, and epiphyseal fusion 
[@Standards1994; @SucheyBrooks1990; @buckberryAuricular2002; @lovejoyAuricular1985; @meindlSutureClosure1985], and divided into the following categories:
early young adult (EYA, 18-24 years), late young adult (LYA, 25-34 years), middle adult (MA, 
35-49 years), old adult (Old, 50+ years).

#### Paleopathology

Pathological conditions and lesions that occur frequently in the population were
included in the analysis. Data were
dichotomised to presence/absence to allow statistical analysis.
Osteoarthritis (OA) was considered present in cases where eburnation was visible
on one or more joint surfaces.
Vertebral osteophytosis is identified by marginal lipping and/or osteophyte
formation on the margin of the superior and inferior surfaces of the vertebral
body.
Cribra orbitalia (CO) was diagnosed based on presence of pitting on the superior
surface of the orbit. No distinction was made between active or healing lesions.
Degenerative disc disease (DDD), or spondylosis, is identified as a large diffuse
depression of the
superior and/or inferior surfaces of the vertebral body [@rogersPalaeopathologyJoint2000].
Schmorl's nodes (SN) are identified as any cortical depressions on the surface of
the vertebral body.
Data on chronic maxillary sinusitis (CMS) from @casnaUrbanizationRespiratory2021 were
included in this study to assess the relationship between upper respiratory
diseases with environmental factors (i.e. tobacco smoke, caffeine consumption).
Lesions associated with CMS as defined by
@boocockMaxillarySinusitis1995 were
recorded for each individual and classified as "pitting", "spicule-type bone
formation", "remodeled spicules", or "white pitted bone". CMS was scored as absent
when the sinus presented smooth surfaces with little or no associated pitting.

#### Dental pathology

Caries ratios were calculated by dividing the number of lesions by the number
of teeth scored, resulting in a single caries ratio per individual. If the surface
where the lesion originated is not visible, i.e. if the lesion covered multiple
surfaces, this was scored as "crown".
Calculus indices were calculated according to Greene and colleagues [-@greeneQuantifyingCalculus2005].
Calculus was scored with a four-stage scoring system (0-3) to score absent, slight,
moderate, and heavy calculus deposits [@brothwellDiggingBones1981]
on the lingual, buccal (labial), and interproximal surfaces of each tooth. Only
one score was used for the combined interproximal surfaces, resulting in three
scores per tooth (when surfaces are intact), and four calculus indices per individual;
upper anterior (UA), upper posterior (UP), lower anterior (LA), lower posterior (LP).
Each index was calculated by dividing
the sum of calculus scores for each surface by the total number of surfaces scored
in each quadrant. If a tooth could not be scored on all three surfaces, the tooth
was not included [@greeneQuantifyingCalculus2005].
Periodontitis was scored on a visual four-stage (0-3) scoring system according to
distance from CEJ of each tooth to alveolar bone [@maatManualPhysical2005].

### Statistical analysis

All compounds and pathological conditions/lesions were converted to a presence/absence
score. Pearson product-moment correlation was applied to the
dichotomised pathological lesions (point-biserial correlation),
compound concentrations, calculus indices, and caries ratios to explore relationships
paired continuous-continuous variables and paired continuous-binary variables.
Compound concentrations were then dichotomised to presence/absence, and
the caries ratio and calculus index for each individual were converted to an
ordinal score from 0 to 4 by using quartiles. Polychoric correlation was applied
to the paired dichotomous variables and dichotomous-ordinal variables.
<!--Median periodontal scores were calculated for each individual. -->

All statistical analysis was conducted in `r version$version.string`,
`r version$nickname`, [@Rbase]. Data wrangling was conducted with the **tidyverse**
[@tidyverse2019] and visualisations were created using **ggplot2** [@ggplot2].
Polychoric correlations were calculated with the **psych** package [@Rpsych].