### 2023-02-07 inst/paper/_discussion.qmd
add co-author comments round 2

<!-- Big summary/important statement for the science -->
In this study we were able to extract and identify multiple alkaloids and salicylic
acid from the
dental calculus of individuals from Middenbeemster, a 19th century Dutch
archaeological site. We applied ultra-high-performance liquid chromatography-tandem
mass spectrometry (UHPLC-MS/MS), a method that was validated by co-occurrence
of drugs and metabolites in dental calculus and blood [@sorensenDrugsCalculus2021].
Here we have shown that the method can also be successfully applied to archaeological
dental calculus.
We extend findings from previous studies on alkaloids in archaeological samples
by extracting multiple different alkaloids from dental calculus, including
nicotine, cotinine, caffeine, theophylline, and salicylic
acid in multiple individuals. The detection of these compounds was
solidified in a replication analysis on different samples from the same
individuals.
Cocaine and multiple cannabinoids were also detected during the first
analysis, but were not replicated.
We discuss the implications of these findings in light of historical and
archaeological evidence for the consumption of these drugs.

<!-- compounds detected with potential sources, and those that were surprisingly missing -->
<!-- nicotine and cotinine: tobacco, obviously...-->
Nicotine and its principal/main metabolite, cotinine, were strongly positively correlated,
both in concentration and presence/absence in individuals (@tbl-pearson and @fig-polycorr).
The detection of nicotine and cotinine is not surprising, as pipe-smoking in the
Beemsterpolder is well-documented in the literature
[@aten400JaarBeemster2012; @boumanBegravenis2017], and visible on the skeletal
remains as pipe notches [@lemmersMiddenbeemster2013].
There is also documented medicinal use of nicotine in the Beemsterpolder,
where a tobacco-smoke enema was used for headaches, respiratory problems, colds,
and drowsiness from around 1780 to 1830 [@aten400JaarBeemster2012].
In our sample, we also detected nicotine and cotinine (replicated)
in an old adult, probable female individual.
In this particular case it is unlikely that the compounds entered the dental
calculus through pipe-smoking, as the individual had no visible pipe notches;
more likely the tobacco entered through an alternate mode of consumption,
secondhand smoke, or the aforementioned tobacco-smoke enema.

<!-- Scientific findings as extension of previous work -->

<!-- strong correlation between nicotine and cotinine concentration is important?
      Could suggest that the incorporation pathway is excretion, not consumption
        - longer time in body than mouth (consumption)
-->

<!-- theophylline and caffeine: coffee and tea? -->
Theophylline and caffeine were positively correlated in our samples, though to a
lesser extent than nicotine and cotinine, so we are unable to determine if they
originated from the same source (@tbl-pearson and @fig-polycorr).
Caffeine and theophylline have very similar chemical structures, so we expect
they would experience similar rates of incorporation and degradation, allowing
us to interpret the ratio and correlations between the compounds.
Caffeine is present in coffee, tea, and cocoa beans, with concentrations slightly
higher in coffee
[@stavricVariabilityCaffeine1988; @chinCaffeineContent2008; @bispoSimultaneousDetermination2002; @srdjenovicSimultaneousHPLC2008].
Theophylline is present in both coffee beans and tea leaves, but in negligible
quantities [@stavricVariabilityCaffeine1988]. It is also a primary metabolite of
caffeine produced by the liver. Given the low correlation, there are likely
multiple sources of caffeine and theophylline in the population, with tea and
coffee being the most obvious.  
Tea consumption had become widespread in the Netherlands by 1820, reaching all
parts of society [@nierstraszTeaTrade2015, p.91].
Historically, we also know that both tea and coffee were consumed in the
Beemsterpolder during the 19th century. 'Theegasten' (teatime) was a special
occasion occurring from 15.00-20.00 hours, where tea was served along with the evening
bread [@schuijtemakerTeTheegasten2011]. Many households also owned at least one
coffee pot and tea pot [@boumanBegravenis2017].
Distinguishing between tea, coffee, and chocolate may be possible by also including
theobromine and comparing ratios of the compounds, as theobromine is present in
higher quantities in chocolate compared to caffeine and theophylline
[@stavricVariabilityCaffeine1988; @bispoSimultaneousDetermination2002; @alanonAssessmentFlavanol2016].
However, In addition to oral factors
affecting alkaloid uptake in dental calculus, there is some indication that
theobromine does not preserve well in the archaeological record
[@velskoDentalCalculus2017], and frequent consumption of all three items would
be difficult to parse.

<!-- salicylic acid: diet, medicine, or environmental??? -->
Salicylic acid was found in all but one individual in our sample.
It can be extracted from the bark of willow trees, *Salix alba*, and
has long been used for its
pain-relieving properties [@bruinsmaBijdragenTot1872, p. 119]. It is also present
in many plant-based foods
[@duthieNaturalSalicylates2011; @malakarNaturallyOccurring2017], including
potatoes, which were a staple of the Beemsterpolder diet [@aten400JaarBeemster2012].
The extracted quantity from our samples decreased over the three washes, followed
by a sharp increase in the final calculus extraction, which is what we would expect
to see if the salicylic acid was incorporated during life [@fig-auth-plot-batch2].
However, it has been shown that salicilyc acid is a very mobile organic acid and
the ubiquitous presence may be due to environmental contamination, which would
also explain the high quantity in the washes
[@chenCa2Dependent2001; @badriRegulationFunction2009].
Given the multiple plausible sources of this residue, it will be necessary to
explore the extent to which salicylic acid can leach
into the dental calculus from the soil, and what the rate of degradation is for
salicylic acid when trapped in dental calculus.
<!-- perhaps using an oral biofilm model ;) -->

<!-- cannabinoids: oh cool! medicinal? wait, where did they go..? -->
Cannabinoids---specifically THC, THCA-A, THCVA, CBD, CBN---were found in the
first batch, but none were replicated in the second batch. Medicinal use of
cannabinoids has been well-established in Europe since Medieval-times, and it
was also grown in the Netherlands [@bruinsmaBijdragenTot1872].
Administration was most common in the form of concoctions containing various portions
of the cannabis plant for ingestion; not until the late 19th century did it become
recommended to smoke it for more immediate effects [@clarkeCannabisEvolution2013].
A Dutch medicinal use of hemp involved an emulsion prepared from the seeds of
the plants to treat pain and various stomach ailments. Another
preparation involving the roots of the plants was used for inflammation, gout,
and joint pains [@clarkeCannabisEvolution2013].
The ability to detect cannabinoids in calculus may be limited by their reduced ability
to diffuse from serum to salivary glands due to an affinity for protein-binding,
[@coneInterpretationOral2007],
meaning detection would rely on oral consumption. Even then,
the overall instability of some cannabinoids could also affect detection
[@sorensenEffectAntioxidants2018; @lindholstLongTerm2010].
However, given the lack of replication, we cannot with security confirm that
cannabis was used by the Beemster population.

<!-- opioids: Damn... that would've been awesome -->
Despite many of our sampled individuals having lived during the height of the opium era
in the Netherlands [@machtHistoryOpium1915], none of the targeted opioids
(morphine, codeine, thebaine, papaverine, norcodeine, noscapine)
were detected. 
The absence of opioids could be a result of the people ascribing more to the
"traditional" rather than "scientific" medicine, although laudanum and
another opium containing concoction was part of the "traditional" medicine in
the Netherlands [@leuwProhibitionLegalizationDutch1994], including Middenbeemster
[@aten400JaarBeemster2012]. It was also generally
considered a drug of the upper class [@scheltemaOpiumTrade1907], and may have been
more common in urban centers.
The absence could also be attributed to postmortem degradation.
It has been shown that, while abundant in opium, morphine
degrades rapidly, while thebaine and papaverine are more resistant to various
ageing processes [@chovanecOpiumMasses2012]. The latter were also absent from
our samples.

<!-- authentication -->
<!-- the importance of replication - brief discussion of cocaine -->
The only strictly modern compound (at least in a European context) detected in the sample
was cocaine, which was detected in the first batch of samples. Our sample is
derived from an early--mid 19th century population, and cocaine was isolated in 1860
by Albert Niemann, and entered popular medical
practice in 1884. Coca arrived in Europe as early as 1771, but as botanical
specimens rather than for consumption, and there were also issues importing enough
viable specimens of coca for cocaine extraction
[@mortimerHistoryCoca1901 p. 179; @abucaCocaTrade2019, p.108].
We considered it possible that it would be present in a sample with
most individuals originating from the early- to mid-19th century. If corroborated,
this would have been the first case of coca-leaf-consumption in Europe.
In our replication batch, we included all
of the individuals who had been cocaine-positive in the first batch. We were
unable to replicate any of the cocaine results, and we were unable to detect the
principal metabolite, benzoylecgonine, in either batch.
We suspect that the original detection of cocaine was a result of lab contamination
during analysis.<!--and hope this will serve as a reminder of the importance of replication.-->

<!-- correlations with skeletal indicators -->
We explored the relationship between detected compounds and various skeletal
indicators, such as pathological and dental lesions, preservation, and pipe notches.
We found some evidence to suggest that preservation of the skeleton influences
the recovery of compounds from the dental calculus, with well-preserved skeletons
potentially serving as a better target for sampling.  
We found a positive correlation between CMS and nicotine, which may be
indicative of the impact tobacco smoking had on the respiratory health of the
Beemster inhabitants.
Tobacco smoke may play a significant role in diseases
of the upper respiratory tract, including chronic maxillary sinusities
[@rehImpactTobacco2012]. Although
the mechanisms by which smoking increases the risk of infections is not fully
understood, solid evidence has been presented linking tobacco smoke to
increased mucosal permeability and impairment of mucociliary clearance
[@arcaviCigaretteSmoking2004]. Such changes, together with an altered immunologic
response, are thought to predispose to the development of chronic maxillary sinusitis
[@slavinDiagnosisManagement2005].  
We also observed a moderate positive correlation between chronic maxillary sinusitis and caffeine which
contradicts previous research linking chronic coffee consumption with a positive
effect on the respiratory system, suggesting a preventive association between
caffeine intake and pneumonia
[e.g. @alfaroChronicCoffee2018; @kondoAssociationCoffee2021].
However, while the
lower respiratory tract seems to benefit from chronic coffee consumption, it is
possible that elevated caffeine intake impacts mucosal moisture due to its
dehydrating effect [@maughanCaffeineIngestion2003], thereby exposing individuals
to greater risk of sinus infection.

<!-- Findings' scientific implications -->

<!-- secondary findings-->

The detection of nicotine in dental calculus has previously been presented by
Eerkens and colleagues [-@eerkensDentalCalculus2018] in two individuals from
pre-contact California. They also targeted caffeine, cotinine, and theophylline
in their samples, but were unable to detect any of them. It remains to be seen
whether this is due to differences in methods used, or due to our samples being
more recent.
They also suggest that the choice of tooth for sampling may impact the detection
of certain compounds,
as the incorporation in dental calculus may depend on the mode of consumption.
Tobacco smokers may have more
nicotine present in calculus on incisors, whereas tobacco chewers may have more
on molars [@eerkensDentalCalculus2018].
However, sampling may not be limited to mode of consumption.
The presence of cotinine suggests that the excretion of a compound after being
metabolised in the body is also a source of deposition, and
that deposition of alkaloids in dental calculus can occur both on the way into
the body, i.e. during consumption, and on the way out, i.e. disposal of waste
products via saliva secretion into the mouth.
Especially mucin-rich saliva from the sublingual
and submandibular glands preferentially binds toxins [@doddsHealthBenefits2005],
and since these glands are located closest to the lower incisors, they may be the
most effective target for these studies.
This has yet to be systematically
tested in archaeological dental calculus.
Because we homogenised samples from multiple teeth of an individual, we were unable
to test the effect of oral biogeography.
It is also possible that resident microflora within
biofilms contribute to alkaloid breakdown and that the presence of caffeine and
nicotine metabolites following direct ingestion can be explained by this pathway.
However, the literature on
biofilm biodegradation of alkaloids is limited, and *in vitro* studies have only
found minimal contributions by certain oral bacteria in isolation
[@sunMetabolomicsEvaluation2016; @cogoVitroEvaluation2008]; it is possible that
a larger role is played by oral bacteria within larger, more metabolically active
communities, e.g. biofilms [@takahashiOralMicrobiome2015].

<!-- How this engineering will facilitate future research -->

<!-- Forward-looking statements about the field as a whole -->

<!-- a summary/importance of the engineering aspect of the work -->

<!-- How it facilitates scientific research -->

<!-- limitations of the platform show which questions would be unanswered -->
Because we targeted individuals with moderate-to-large calculus deposits,
it is likely a biased sample. The presence of calculus may increase the
risk of premature death
[@yaussyCalculusSurvivorship2019], and periodontal disease (which may or may not
be associated with dental calculus build-up) is a risk-factor for
respiratory diseases, if periodontal and respiratory pathogens
enter the bloodstream
[@azarpazhoohSystematicReview2006; @scannapiecoRoleOral1999; @scannapiecoPotentialAssociations2001].
In our sample, the percentage of chronic maxillary sinusitis
(`r scales::percent(mean(explore_select$CMS, na.rm = T), accuracy = 0.1)`)
is lower than in another (more representative) male sample (44.1%)
[@casnaUrbanizationRespiratory2021], and the caries percentage is similarly lower
in our sample (`r scales::percent(caries_ratio_site$ratio, accuracy = 0.1)`) than
a more representative sample (22.9%) [@lemmersMiddenbeemster2013].  
We used the presence/absence of a pipe notch and concurrent detection of tobacco
as a crude estimate of the accuracy of the method, which we found to be around
`r scales::percent(mean(replicated_accuracy$correct, na.rm = T), accuracy = 0.1)`.
This is a very rough estimate, as the presence of a pipe notch is likely not a
perfect indicator of whether or not someone consumed tobacco.
Dental calculus is also
more transient than for example bone, as it can be mechanically removed,
intentionally or unintentionally, during life, eliminating
all trace of the alkaloids consumed prior to its removal.  
Quantitation of the detected compounds may have limited value in archaeological
samples due to degradation, and will greatly affect our correlations related to
concentration.
Following burial, compound stability over
time will play a large role, as will microbial degradation of compounds by bacteria
and fungi in soil [@liuNicotinedegradingMicroorganisms2015], as well as
the soil environment, such as temperature, pH, and oxygen availability
[@mackiePreservationMetaproteome2017; @lindholstLongTerm2010].  
The detected quantity of a compound will also depend on the quantity in dental
calculus during life, which is largely controlled by quantity of consumption,
how often the
calculus was disrupted/removed, metabolic breakdown of the compound, and inter-
and intra-individual factors related to stages of biofilm formation, maturation,
and mineralisation
[@lustmannScanningElectron1976; @velskoMicrobialDifferences2019; @zijngeBiofilmArchitecture2010].
In short, this means it is not really possible to detect the absence of a compound.
The absence of a compound is not evidence of absence of consumption. This complicates
the interpretation of our results.
We have attempted to minimise errors occurring due to this limitation by including
a relatively large sample of individuals and replicating our analysis. Although
given the relatively low detection rate seen in tobacco, this remains a major
limitation, and will likely be compounded by increasing antiquity of the samples.

Future studies should explore how sampling from various types of teeth
and their position in the mouth affects the probability of a compound becoming
entrapped in dental calculus.
This may also be related to properties within the oral cavity, as well as chemical
properties of the compounds, which facilitate or reduce the incorporation-potential,
and which incorporation pathways are more likely for a given compound.  
We only targeted drugs that were included in the forensic toxicological
screenings, and therefore only covered a limited number of the potential compounds
that could be of interest for exploring past diets and medicinal treatments. The
list of targeted compounds can be expanded as we discover more potential targets
based on which specific compounds/metabolites are more likely to be incorporated
and preserved in dental calculus.  
There is an increasing interest in using oral fluid as a means of detecting alkaloids
in living
individuals due to the non-invasive nature of the testing compared to blood
and urine sampling [@coneSalivaTesting1993; @valenDetermination212017].
These *in vivo* studies are a valuable
source of method validation and can help determine the feasibility of detecting
certain alkaloids in oral fluid and, subsequently, dental calculus. Archaeologists,
though, will likely be responsible for exploring dental calculus specific
incorporation and retention of alkaloids, as well as their long-term preservation
in the burial environment.<!-- move to discussion in disseration: Here, *in vitro* studies may provide valuable insights-->

<!-- How it will facilitate the specific topic of research (e.g. archaeology) -->

<!-- rephrase -->
While a major limitation is the uncertainty surrounding whether or not a compound
is actually absent, the power of the method lies
in the ability to detect dietary and other compounds that were incorporated
via multiple
consumption pathways that are not detected by other methods.
Taking
tobacco consumption as an example; while pipe
notches are a useful way to identify tobacco consumption, pipe smoking was not the
only mode of tobacco consumption, with others including chewing, drinking, cigars,
and snuff [@goodmanTobaccoHistory1994, p.67]. Pipe-smoking
was mainly practised by males [@lemmersMiddenbeemster2013; @eerkensDentalCalculus2018],
so methods like the one presented here
are suitable for exploring tobacco consumption in an entire society, rather
than a trivial subset of past populations.<!--admittedly, our sample was just that...-->
Combined with other methods, it can also give us a more complete picture
of dietary patterns and medicinal/recreational plant-use in the past by capturing
multiple possible incorporation pathways of dietary (and other) compounds.