### 2023-02-07 analysis/supplementary-materials/supp-mat.qmd

---
title: "Supplementary Materials"
execute: 
  echo: false
  warning: false
knitr: 
  opts_chunk: 
    message: false
format: 
  html:
    toc: true
    code-fold: true
---

```{r}
#| label: setup
#| include: false
library(here)
library(readr)
devtools::load_all()

# upload data
metadata <- read_tsv(here("inst/data/raw_data/metadata.tsv"))
demography <- read_csv(here("inst/data/raw_data/demography.csv"))
lloq <- read_tsv(here("inst/data/raw_data/lloq.tsv"))
uhplc_data_comb <- read_csv(here("inst/data/derived_data/uhplc-data_combined.csv"))
#uhplc_calculus <- read_csv(here("inst/data/derived_data/uhplc-calculus_cleaned.csv"))
dental_inv <- read_csv(here("inst/data/raw_data/dental-inv.csv"))
caries <- read_csv(here("inst/data/raw_data/caries.csv"))
periodont <- read_csv(here("inst/data/raw_data/periodontitis.csv"))
periap <- read_csv(here("inst/data/raw_data/periapical.csv"))
calculus <- read_csv(here("inst/data/raw_data/calculus.csv"))
calculus_full <- read_csv(here("inst/data/raw_data/calculus_full.csv"))
sinusitis_clean <- read_csv(here("inst/data/derived_data/sinusitis_cleaned.csv"))
path_cond_clean <- read_csv(here("inst/data/derived_data/path-conditions_cleaned.csv"))

source(here("inst/scripts/setup-qmd.R"))
```

## Session information

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r}
print(sessionInfo(), locale = F)
```

## Data dictionary

### metadata

<!-- id	sample	element	KZ_sample	KZ_element	replicated	batch1_weight	batch2_weight -->

| variable | description |
|---|---|
| id | unique identifier for the individual |
| sample | sample number for the UHPLC-MS/MS analysis |
| element | which tooth (FDI notation) was sampled for UHPLC-MS/MS analysis |
| replicated | Whether the sample was included in the replication batch (TRUE/FALSE) |
| batch1_weight | weight (mg) of sample in batch 1 |
| batch2_weight | weight (mg) of sample in batch 2 |

### dental-inv

Dental inventory

| variable | description |
|---|---|
| id | unique identifier for the individual |
| t11..t48 | Tooth number (FDI system) and status |

Dental inventory key:

- p = present
- m = missing (for unknown reason)
- aml = ante-mortem loss
- dna = removed for DNA sampling

### caries

Caries lesions location

| variable | description |
|---|---|
| id | unique identifier for the individual |
| t11..t48 | Tooth number (FDI system) and location of caries |

Caries location key:

- none = no caries present
- mes = mesial
- dis = distal
- lin = lingual (and labial)
- buc = buccal
- occ = occlusal
- crown = crown (caries-related loss of much of/entire tooth crown)
- root = root
- blank = tooth not present

### periodontitis

| variable | description |
|---|---|
| id | unique identifier for the individual |
| t11..t48 | Tooth number (FDI system) and periodontitis score (0-3) |

Periodontitis score:

- 0 = none
- 1 = slight
- 2 = moderate
- 3 = severe

- blank = tooth not present

### periapical

| variable | description |
|---|---|
| id | unique identifier for the individual |
| t11..t48 | Tooth number (FDI system) and periapical lesion |

- none = no lesion
- bucc = buccal
- lin = lingual
- perf = perforated alveolar bone
- blank = not scoreable

### calculus

| variable | description |
|---|---|
| id | unique identifier for the individual |
| t11..t48 | Tooth number (FDI system) and deposit size (0-3) |

- 0 = no calculus
- 1 = slight calculus
- 2 = moderate calculus
- 3 = heavy calculus

Scored according to Standards.

### path-conditions

Pathological conditions

- OA = osteoarthritis
- IVDD = intervertebral disc disease
- TB = tuberculosis
- DISH = diffuse idiopathic skeletal hyperostosis
- VOP = vertebral osteophytosis
- SN = schmorl's nodes
- DDD = degenerative disc disease
- PNBF = periosteal new bone formation
- OD = osteochondritis dissecans
- CF = cribra femora
- CO = cribra orbitalia

## Sample

### Selection

```{r}
#| label: tbl-sampling
#| tbl-cap: "Table showing which individuals were sampled in this study (element) and which individuals were sampled in the previous study (KZ_element)."
metadata %>% 
  select(id, element, KZ_element) %>% 
  mutate(
    element = case_when(is.na(element) ~ FALSE,
                   TRUE ~ TRUE),
    KZ_element = case_when(is.na(KZ_element) ~ FALSE,
                   TRUE ~ TRUE),
    )
```

### Demographics

```{r}
#| label: male-sample
males <- demography %>%
  filter(sex == "m" | sex == "pm")
```

The sample consists of `r nrow(demography)` individuals, most of which are
middle adult male individuals (@fig-age-distribution).

```{r}
#| label: fig-age-distribution
demography %>%
  ggplot(aes(x = age, fill = sex)) +
    geom_bar() +
    theme_minimal()
```

### Missing data

Missing data per dataset

```{r}
#| label: fig-missing-caries
caries %>%
  mutate(across(!id, ~ is.na(.x))) %>%
  select(!id) %>%
  colSums()

caries %>%
  mutate(across(!id, ~ is.na(.x))) %>% 
  pivot_longer(-id) %>% 
  ggplot(aes(x = name, y = id, fill = value)) +
    geom_tile()
```

Missing data across all dental datasets

```{r}
#| label: fig-missing-dental
#| fig-cap: "Number of missing values per individual and tooth across all dental scores (caries, periodontitis, periapical lesions, calculus)."
dental_long %>%
  mutate(across(c(caries, periodont, periap, calculus), is.na)) %>%
  select(c(caries, periodont, periap, calculus, id, tooth)) %>%
  #group_by(id, tooth) %>% 
  rowwise() %>% 
  mutate(missing = sum(caries, periodont, periap, calculus)) %>% 
  ggplot(aes(x = tooth, y = id, fill = missing)) +
    geom_tile()
```

## UHPLC analysis

```{r}
#| label: setup-uhplc

# presence/absence data frame
uhplc_calculus_bin <- uhplc_calculus_long %>%
  mutate(presence = if_else(quant > 0, 1, quant))

# successfully replicated samples only
uhplc_calculus_replicated <- uhplc_calculus_bin %>%
  mutate(compound = str_remove(compound, "_calc")) %>%
  group_by(id, sample, compound) %>% # combine batches
  summarise(presence = sum(presence)) #%>%
  # filter(presence == 0 | presence == 2) %>%
  # group_by(compound) %>%
  # mutate(presence = if_else(presence == 2, 1, 0)) # convert replications to presence/absence

uhplc_calculus_replicated <- uhplc_calculus_bin %>%
  filter(id %in% filter(metadata, replicated == TRUE)$id) %>%
  #right_join(select(filter(metadata, replicated == TRUE)), by = c("id", "sample"))
  #mutate(compound = str_remove(compound, "_calc")) %>%
  group_by(id, sample, compound) %>% # combine batches
  summarise(presence = sum(presence)) %>%
  filter(presence == 0 | presence == 2) %>% # remove compounds only detected in one batch
  group_by(compound) %>%
  mutate(presence = if_else(presence == 0, 0, 1)) %>%  # convert replications to presence/absence
  ungroup()

uhplc_replicated_wide <- uhplc_calculus_replicated %>%
  mutate(compound = case_when(compound == "nicotine" ~ "tobacco",
                              compound == "cotinine" ~ "tobacco",
                              TRUE ~ compound)) %>%
  group_by(id, sample, compound) %>%
  summarise(presence = sum(presence)) %>% # combine nicotine and cotinine
  #remove_missing() %>%
  mutate(presence = case_when(presence > 0 ~ TRUE,
                              TRUE ~ FALSE)) %>%
  pivot_wider(names_from = "compound", values_from = "presence")
```

The UHPLC-MS/MS method was validated in a separate study on cadavers received
for forensic autopsy and toxicological analysis. Results from dental calculus were
validated against compounds detected in whole blood samples from the same
individuals [@sorensenDrugsCalculus2021].  
In the original method, samples were washed three times to remove residual
substances from the surface of the calculus that originated from oral fluids,
and only extract substances from the calculus. In our samples the washes served
to remove potential contaminants from the burial environment and post-excavation
handling.

Briefly, dental calculus was treated with citric acid and the dissolution extracts
were cleaned using weak and strong polymeric cation-exchange sorbents. Samples
were washed with 0.5 mL MeOH for 10 seconds. Samples were weighed before and after
each wash. The wash solvent was evaporated to a residual volume of 10 µl and added
50µl 30% methanol.
Samples were air-dried for 24 hours at room temperature after each wash.
Extracts from each wash were analysed by injecting 5 µL into the column
on an Exion UHPLC system that consisted of two Exion AD pumps, an Exion AD
multiplate autosampler set at 10 $\pm$ 2 &deg;C and an Exion AC column oven set
at 40 $\pm$ 2 &deg;C (Sciex, Ontario, Canada).
Separation was performed using a Raptor Biphenyl UHPLC column (2.7 mm, 2.1 mm I.D.
$\times$ 100 mm) (Restek, Bellefonte, PA). The mass spectrometer was a Sciex QTRAP
6500+ with a TurboIonSpray probe for electrospray ionisation.  
The remaining calculus was dissolved using lysing tube beads in 800 $\mu$L of 0.5
$\small{M}$ citric acid (CA) and 50 $\mu$L stable isotope-labelled analogue used
as internal standards (SIL-IS) solution for 1 h at ambient
temperature with gentle shaking. The suspension was then mixed with 800 $\mu$L
MeOH and centrifuged at 10,000 $\times$ g for 5 mins, and analysed by the same
method as the wash extracts.
Data analysis was performed using Analyst 1.7 and MultiQuant 3.0.3 (Sciex).
Raw quantities of compounds are presented in ng and concentrations as ng / mg.

The samples in the replication batch were processed in the same way, but
analysed on equipment used exclusively for oral samples.

```{r}
#| label: tbl-lloq
#| tbl-cap: "Target compounds and lower limits of quantitation."
lloq %>%
  arrange(compound)
```

```{r}
#| label: tbl-uhplc-batch-1
#| tbl-cap: "Results from the UHPLC analysis first batch. Quantity of compound in the dissolved calculus, represented in ng."
uhplc_calculus_long %>%
  filter(batch == "batch1") %>%
  #select(sample, contains("batch1")) %>%
  #rename_with(~ str_remove(.x, "_calc_batch1"), .cols = contains("_calc_batch1"))
  pivot_wider(names_from = "compound", values_from = "quant") %>%
  select(!c(extraction, batch))
```

```{r}
#| label: tbl-uhplc-batch-2
#| tbl-cap: "Results from the UHPLC analysis second batch. Quantity of compound in calculus after third wash, represented in ng."
uhplc_calculus_long %>%
  filter(batch == "batch2") %>%
  #select(sample, contains("batch1")) %>%
  #rename_with(~ str_remove(.x, "_calc_batch1"), .cols = contains("_calc_batch1"))
  pivot_wider(names_from = "compound", values_from = "quant") %>%
  select(!c(extraction, batch))
```

### Authentication

No modern synthetic drugs were detected in any of the samples.

Samples were replicated to verify results from the initial analysis. Of the
`r nrow(demography)` samples initially analysed,
`r nrow(filter(metadata, replicated == TRUE))` samples were replicated.

Only caffeine, theophylline, nicotine, cotinine, and salicylic acid were found
in the replicated samples.

```{r}
#| label: tbl-uhplc-replicated
uhplc_calculus_replicated %>%
  filter(sum(presence) != 0) %>% # remove compounds absent in all individuals in sample
  pivot_wider(names_from = "compound", values_from = "presence")
```

<!-- line plot with concentration across washes and to the final calculus sample -->

Most plots show a large increase in extracted mass of a compound between the
calculus wash extracts (wash 1-3) and the dissolved calculus (calc). Most samples
containing theophylline and caffeine had the largest quantity of the compound
extracted from the first wash, then decreasing in washes 2 and 3. There is
an increase between wash 3 and the dissolved calculus in all samples.
The patterns are consistent across batches 1 and 2. The pattern we expect to see
in a sample is a reduction in the quantity from wash 1 to wash 3, and then another
spike in the final extraction from the dissolved calculus, which means the compound
is actually 'ancient' or authentic. The compounds that are completely absent in
all three washes and present in high quantities in the final extraction may also
be suggestive of lab contamination.
This has not been thoroughly tested and is
only based on what we expect to see. Therefore, the interpretation of these graphs
is itself up for interpretation.

```{r auth-plot-batch1}
#| label: fig-auth-plot-batch1

uhplc_data_long %>%
  filter(
    batch == "batch1",
    ) %>% 
  semi_join(quant_filter, by = c("sample", "compound")) %>% 
  mutate(
    extraction = factor(extraction, levels = c("wash1", "wash2", "wash3", "calc")),
    sample = as.factor(sample)
    ) %>% 
  ggplot(aes(x = extraction, y = quant, group = sample, colour = sample)) +
    geom_line() +
    geom_point(size = 0.2) +
    facet_wrap(~ compound, scales = "free_y") +
    theme_bw()
```

```{r}
#| label: fig-auth-plot-batch2
uhplc_data_long %>%
  filter(
    batch == "batch2",
    !compound %in% c("cbd", "cbn", "cocaine", "thc", "thca-a", "thcva") # remove compounds not detected in batch 2
    ) %>% 
  semi_join(quant_filter, by = c("sample", "compound")) %>% # remove compounds not detected in each sample 
  mutate(
    extraction = factor(extraction, levels = c("wash1", "wash2", "wash3", "calc")),
    sample = as.factor(sample)
    ) %>% 
  ggplot(aes(x = extraction, y = quant, group = sample, colour = sample)) +
    geom_line() +
    geom_point(size = 0.2) +
    facet_wrap(~ compound, scales = "free_y") +
    theme_bw()
```

### Quantity vs. sample weight

There is no clear relationship between the sample weight and the amount of compound
detected, except for salicylic acid, where the amount of extracted compound increases
with increasing sample weight. In batch 2 there is also a slight positive trend
for caffeine, nicotine, and cotinine.
Nicotine and cotinine display the same relative relationship between samples. Where
the nicotine quantity is high compared to other samples, the cotinine quantity
will be similarly high (@fig-quant-weight-1 and @fig-quant-weight-1).

The positive correlation between the weight of the calculus
sample and recovered quantities of the compounds suggests sample weight may affect
the ability to detect compounds; although, we were able to detect
compounds in samples as small as 2 mg [@fig-quant-weight-1 and @fig-quant-weight-2].

```{r}
#| label: fig-quant-weight-1
#| #| fig-cap: "Quantity of a compound (ng) found in a sample plotted against the weight of the calculus sample. Results from batch 1."
uhplc_data_comb %>%

  select(sample, contains("batch1")) %>% 
  select(sample, batch1_weight, contains("calc")) %>% 
  pivot_longer(
    -c(sample, batch1_weight),
    names_to = c("compound", "batch"),
    names_pattern = "(.*)_(.*)",
    values_to = c("conc")
  ) %>% 
  filter(conc > 0) %>% 
  mutate(compound = str_remove(compound, "_calc")) %>%
  ggplot(aes(x = batch1_weight, y = conc, col = as.factor(sample))) +
    geom_point() +
    facet_wrap(~ compound, scales = "free_y") +
    theme_bw() +
    theme(legend.position = "none") +
    labs(x = "", y = "Quantity (ng)")
```

```{r}
#| label: fig-quant-weight-2
#| fig-cap: "Quantity of a compound (ng) found in a sample plotted against the weight of the calculus sample. Results from batch 2."
uhplc_data_comb %>%
  select(sample, contains("batch2")) %>% 
  select(sample, batch2_weight, contains("calc")) %>% 
  pivot_longer(
    -c(sample, batch2_weight),
    names_to = c("compound", "batch"),
    names_pattern = "(.*)_(.*)",
    values_to = c("conc")
  ) %>% 
  mutate(compound = str_remove(compound, "_calc")) %>%
  #remove_missing() %>% 
  filter(conc > 0) %>% 
  ggplot(aes(x = batch2_weight, y = conc, col = as.factor(sample))) +
    geom_point() +
    facet_wrap(~ compound, scales = "free_y") +
    theme_bw() +
    labs(y = "Quantity (ng)", y = "", col = "Sample number")
```

### Distribution of compounds detected in the samples

<!-- absolute counts in each batch -->

```{r}
#| label: fig-compounds-detect
#| fig-cap: "Number of individuals in which each compound was detected between batch 1 and 2."

# awkward wrangling required because the information on which individuals were included in replication was lost when combining batches 1 and 2
#n_batch1 <- length(metadata$replicated)
#n_batch2 <- sum(metadata$replicated) # sum boolean vector to get length of TRUE values
uhplc_calculus_long %>%
  #remove_missing() %>%
  filter(quant > 0) %>% 
  #mutate(detect = replace_na(conc, 0)) %>% 
  #mutate(detect = if_else(detect > 0, 1, 0)) %>%
  #group_by(compound, batch) %>% 
  # summarise(
  #   n_detect = sum(detect)
  #           ) %>% 
  # mutate(n = case_when(batch == 1 ~ n_batch1,
  #                      batch == 2 ~ n_batch2)) %>% 
  # mutate(prop = n_detect / n) %>%

ggplot(aes(x = compound, fill = compound)) +
    #geom_jitter(width = 0.2) +
    #geom_boxplot() +
    geom_bar() +
    facet_wrap(~ batch) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
      legend.position = "none"
      )
```

The replication showed that caffeine, theophylline, cotinine, nicotine, and
salicylic acid could be consistently detected in the samples, although theophylline
detection decreased between batches 1 and 2. CBD, CBN, cocaine, and THCA-A was not
detected at all in the second batch.

<!-- absolute counts in the replicated individuals -->

```{r}
#| label: fig-compounds-detect2
#| fig-cap: "Number of individuals in which each compound was detected between batch 1 and 2. Only showing replicated individuals."

uhplc_calculus_long %>%
  filter(id %in% filter(metadata, replicated == T)$id,
         quant > 0) %>%
  #remove_missing() %>%
  ggplot(aes(x = compound, fill = compound)) +
    geom_bar() +
    facet_wrap(~ batch) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
      legend.position = "none"
      )
```

### Detection and preservation

To see if preservation of the skeletal remains had any effect on the detection of
compounds, absolute quantities of compounds were compared to the various levels of
preservation.

```{r}
#| label: fig-detection-preservation
uhplc_calculus_long %>%
  filter(
    !is.na(preservation),
    batch == "batch1",
    quant > 0,
    #!compound %in% c("cbd", "cbn", "cocaine", "thc", "thca-a", "thcva"), # remove compounds not detected
    ) %>% 
  ggplot(aes(x = preservation, y = quant)) +
    geom_violin(aes(fill = preservation), alpha = 0.6) +
    #geom_jitter(width = 0.2) +
    geom_boxplot(width = 0.2) +
    facet_wrap(~ compound, scales = "free_y", labeller = labeller(compound = compound_names), dir = "v") +
    theme_bw() +
    theme(
      legend.position = "none",
      panel.grid.major.x = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
      ) +
    labs(x = "Preservation", y = "Quantity (ng)") +
    scale_fill_viridis_d()
```

```{r}
#| label: fig-detection-preservation2
uhplc_calculus_long %>%
  filter(
    !is.na(preservation),
    quant > 0,
    batch == "batch2",
    #!compound %in% c("cbd", "cbn", "cocaine", "thc", "thca-a", "thcva"), # remove compounds not detected
    ) %>% 
  ggplot(aes(x = preservation, y = quant)) +
    geom_violin(aes(fill = preservation), alpha = 0.6) +
    #geom_jitter(width = 0.2) +
    geom_boxplot(width = 0.2) +
    facet_wrap(~ compound, scales = "free_y", labeller = labeller(compound = compound_names), dir = "v") +
    theme_bw() +
    theme(
      legend.position = "none",
      panel.grid.major.x = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
      ) +
    labs(x = "Preservation", y = "Quantity (ng)") +
    scale_fill_viridis_d()

```

Distribution of state-of-preservation in batches 1 and 2 to make sure the number
of skeletons are not affecting the relationships shown above. Given our sample
contains a smaller number of individuals with fair preservation, this may bias
our interpretations (@fig-preservation-detection).

```{r}
#| label: fig-preservation-detection
uhplc_calculus_long %>% 
  ungroup() %>% 
  distinct(id, batch, .keep_all = TRUE) %>%
  remove_missing(vars = c("preservation", "weight")) %>% 
  ggplot(aes(x = preservation)) +
    geom_bar() +
    facet_wrap(~ batch) +
    theme_bw()
```

### Detection of tobacco

Given that pipe notches are present in the majority of individuals, the presence
of pipe notch(es) in an individual and concurrent detection of nicotine and/or
cotinine is used as a rough indicator of the accuracy of the method.

```{r}
#| label: fig-corr-notches
# correlation between nicotine and cotinine conc, and pipe notches
uhplc_calculus_long %>%
  ggplot(aes(x = pipe_notch, y = conc)) +
    geom_point() +
    facet_wrap(~ compound + batch, scales = "free_y")
```

```{r}
tobacco <- uhplc_calculus_long %>%
  filter(compound %in% c("nicotine", "cotinine")) %>% 
  mutate(detection = if_else(is.na(conc), FALSE, TRUE))

tobacco <- uhplc_calculus_long %>%
  filter(
    #batch == "batch2",
    compound %in% c("nicotine", "cotinine"),
    sex == "m" | sex == "pm"
    ) %>% 
  mutate(detection = if_else(quant == 0, 0, 1)) # 0 = not detected; NA = not included in batch 2.

tobacco_accuracy <- tobacco %>%
  remove_missing(vars = "quant") %>%
  group_by(sample, id, .drop = F) %>% 
  summarise(
    detection = sum(detection),
    #compound = compound
    ) %>% 
  left_join(select(demography, id, pipe_notch, preservation, age), by = "id") %>% 
  mutate(
    pipe_notch = if_else(pipe_notch > 0, "Y", "N"),
    correct = case_when(
      detection == 0 & pipe_notch == "N" ~ 1,
      detection > 0 & pipe_notch == "Y" ~ 1,
      detection > 0 & pipe_notch == "N" ~ NaN, # no way of knowing macroscopically if the person smoked without pipe
      TRUE ~ 0
    )
  )

accuracy_age <- tobacco_accuracy %>% # move to supplementary material
  group_by(age) %>% 
  summarise(mean = mean(correct, na.rm = T))

tobacco_comb <- tobacco %>% 
  group_by(batch, sample, id, .add = TRUE) %>%
  summarise(
    detection = sum(detection), 
    #sex, age, completeness, preservation, pipe_notch, 
    .groups = "keep" # combine compounds: 0 = none detected; 1 = 1 detected; 2 = both detected
  ) %>%
  ungroup() %>%
  left_join(select(demography, id, pipe_notch), by = "id") %>%
  mutate(
    pipe_notch = if_else(pipe_notch > 0, "Y", "N"),
    correct = case_when(
      detection > 0 & pipe_notch == "Y" ~ TRUE,
      detection == 0 & pipe_notch == "N" ~ TRUE,
      detection > 0 & pipe_notch == "N" ~ NA, # can't be sure if this is true or not
      TRUE ~ FALSE
      )
    )

nicotine <- tobacco %>%
  filter(compound == "nicotine") %>% 
  mutate(
    pipe_notch = if_else(pipe_notch > 0, "Y", "N"),
    correct = case_when(
      detection > 0 & pipe_notch == "Y" ~ TRUE,
      detection == 0 & pipe_notch == "N" ~ TRUE,
      detection > 0 & pipe_notch == "N" ~ NA, # can't be sure if this is true or not
      TRUE ~ FALSE
    )
  )

cotinine <- tobacco %>%
  filter(compound == "cotinine") %>% 
  mutate(
  correct = case_when(
    detection > 0 & pipe_notch == "Y" ~ TRUE,
    detection == 0 & pipe_notch == "N" ~ TRUE,
    detection > 0 & pipe_notch == "N" ~ NA, # can't be sure if this is true or not
    TRUE ~ FALSE
    )
  )

# accuracy in replicated samples
uhplc_accuracy <- uhplc_calculus_replicated %>%
  left_join(select(demography, id, pipe_notch), by = "id") %>% 
  filter(compound == "nicotine" | compound == "cotinine") %>%
  group_by(id) %>% 
  arrange(desc(presence)) %>%
  slice(which(compound == "nicotine" || compound == "cotinine")) %>%
  mutate(
    pipe_notch = if_else(pipe_notch > 0, "Y", "N"),
    correct = case_when(presence == 1 & pipe_notch == "Y" ~ TRUE,
                             presence == 0 & pipe_notch == "N" ~ TRUE, # can we be sure if this is correct?
                             presence == 1 & pipe_notch == "N" ~ NA, # can't be sure if this is true or not
                             TRUE ~ FALSE))

# ratio of nicotine to cotinine
tobacco_ratio <- tobacco %>%
  select(id, batch, compound, conc) %>% 
  pivot_wider(names_from = compound, values_from = conc) %>% 
  group_by(id, batch) %>% 
  summarise(ratio = cotinine / nicotine) %>%
  remove_missing() %>% 
  left_join(select(demography, id, age, sex, preservation, pipe_notch)) %>% 
  filter(batch == "batch2",
         ratio != Inf) %>% 
  mutate(
    age = case_when(
      age == "eya" ~ 0,
      age == "lya" ~ 1,
      age == "ma" ~ 2,
      age == "old" ~ 3
    ),
    preservation = case_when(
      preservation == "fair" ~ 0,
      preservation == "good" ~ 1,
      preservation == "excellent" ~ 2
    )
  )

tobacco_ratio %>%
  ggplot(aes(y = ratio, x = pipe_notch)) +
    #geom_boxplot()
    geom_point()

  #lm(ratio ~ age + sex + preservation + pipe_notch, tobacco_ratio)
```

We found no correlation between the number of pipe notches
and the concentration of nicotine and cotinine, suggesting that our ability
to detect tobacco consumption in dental calculus does not necessarily rely on
targeting frequent smokers; here, we consider individuals with multiple pipe notches
as likely to have been heavy smokers.

No apparent correlation between the number of pipe notches
and the concentration of nicotine or cotinine [@fig-corr-notches].

The presence of pipe notch(es) in an individual and concurrent detection of nicotine
and/or cotinine is used as a crude indicator of the accuracy of the method. When
combining the results of both batches, the method was able to detect some form
of tobacco in
`r nrow(filter(tobacco_accuracy, pipe_notch == "Y", detection > 0))`
of `r nrow(filter(tobacco_accuracy, pipe_notch == "Y"))`
individuals with a pipe notch
(`r scales::percent(nrow(filter(tobacco_accuracy, pipe_notch == "Y", detection > 0)) / nrow(filter(tobacco_accuracy, pipe_notch == "Y")), accuracy = 0.1)`).
When also considering correct the absence of a tobacco alkaloid together with the absence
of a pipe notch, the accuracy of the method is
`r scales::percent(mean(tobacco_accuracy$correct, na.rm = T), accuracy = 0.1)`.
Accuracy in the old adult age category is
`r scales::percent(filter(accuracy_age, age == "old")$mean, accuracy = 0.1)`.

In the replicated samples only, tobacco detection was successful in
`r nrow(filter(replicated_accuracy, pipe_notch == "Y", detection > 0))`
out of
`r nrow(filter(replicated_accuracy, pipe_notch == "Y"))`
pipe smokers
(`r scales::percent(nrow(filter(replicated_accuracy, pipe_notch == "Y", detection > 0)) / nrow(filter(replicated_accuracy, pipe_notch == "Y")), accuracy = 0.1)`)
Including individuals with absence of a pipe notch and concurrent absence of
compounds as a correct identification, gives an overall accuracy of
`r scales::percent(mean(replicated_accuracy$correct, na.rm = T), accuracy = 0.1)`.

Nicotine and cotinine detection successfully identified
`r nrow(filter(tobacco_comb, batch == 1, pipe_notch == "Y", detection > 0))`
smokers out of
`r nrow(filter(tobacco_comb, batch == 1, pipe_notch == "Y"))`
(`r scales::percent(nrow(filter(tobacco_comb, pipe_notch == "Y", batch == 1, detection > 0)) / nrow(filter(tobacco_comb, batch == 1, pipe_notch == "Y")), accuracy = 0.1)`)
individuals with pipe notches in the first analysis.

Nicotine detection successfully identified
`r nrow(filter(nicotine, batch == 1, pipe_notch == "Y", detection > 0))`
smokers out of
`r nrow(filter(nicotine, batch == 1, pipe_notch == "Y"))`
individuals with pipe notches
(`r scales::percent(nrow(filter(nicotine, batch == 1, pipe_notch == "Y", detection > 0)) / nrow(filter(nicotine, batch == 1, pipe_notch == "Y")), accuracy = 0.1)`).

Cotinine detection successfully identified
`r nrow(filter(cotinine, batch == 1, pipe_notch == "Y", detection > 0))`
smokers out of
`r nrow(filter(cotinine, batch == 1, pipe_notch == "Y"))`
individuals with pipe notches
(`r scales::percent(nrow(filter(cotinine, batch == 1, pipe_notch == "Y", detection > 0)) / nrow(filter(cotinine, batch == 1, pipe_notch == "Y")), accuracy = 0.1)`).

In the replicated samples only, tobacco detection was successful in
`r nrow(filter(uhplc_accuracy, pipe_notch == "Y", presence > 0))`
out of
`r nrow(filter(uhplc_accuracy, pipe_notch == "Y"))`
smokers
(`r scales::percent(mean(filter(uhplc_accuracy, pipe_notch == "Y")$correct, na.rm = T))`).
Including individuals with absence of a pipe notch and concurrent absence of
compounds as a correct identification, gives an overall accuracy of
`r scales::percent(mean(uhplc_accuracy$correct, na.rm = T), accuracy = 0.1)`.

One individual---an old adult, probable female---was positive
for both nicotine and cotinine, and had no signs of a pipe notch.

## Dental analysis

Pipe notches were identified by wear on the mesial and distal
sides of the crowns between to teeth, resulting from the practice of clenching a
pipe between adjacent and isomeric teeth, and which differs from the occlusal
wear that occurs through mastication. Wear occurring between adjacent and isomeric
teeth were counted as a single pipe notch.

```{r}
#| label: fig-dental-plots
#| layout-ncol: 2
#| fig-cap: 
#|   - "Antemortem tooth loss in sample."
#|   - "Periodontal scores in the sample."
dental_long %>%
  dental_plot(fill = status)

dental_long %>%
  filter(!is.na(periodont)) %>% 
  dental_plot(fill = periodont)
```

### AMTL

### Caries

Caries were scored as the location on each individual tooth. Multiple locations on
a single tooth were separated with `;`. The size of caries was also
recorded, but not used in further analysis. Large caries that cover multiple
surfaces with an unknown origin were recorded as 'crown'. Caries counts were
extracted based on the number of locations recorded.

| code | surface |
|---|---|
| mes | mesial surface |
| dis | distal surface |
| lin | lingual surface |
| buc | buccal surface (including labial surface) |
| occ | occlusal surface (including incisal surface) |
| crown | caries covers 2+ surfaces |
| none | No caries visible on surface |
| NA | Not observable/tooth missing |

```{r}
surface <- c("mes", "dis", "occ", "buc", "lin", "root", "crown") # here occ = incisal

# caries_count <- dental_long %>% 
#   select(!c(periodont, periap, calculus)) %>% 
#   #filter(status == "p") %>%
#   separate_rows(caries, sep = ";") %>% # one lesion per row
#   mutate(caries_count = if_else(
#     caries == "none", 0L, 1L
#     )
#   ) %>%
#   group_by(id, tooth) %>%
#   summarise(
#     caries_count = sum(caries_count, na.rm = F)
#     ) %>%
#   tooth_position()
# 
# site_wide_caries <- caries_count %>%
#   caries_ratio(.caries = caries_count)
```

In the `r caries_ratio_site$n_teeth` that were examined, `r caries_ratio_site$count`
teeth had caries
(`r scales::percent(caries_ratio_site$ratio, accuracy = 0.1)`).
This frequency has very little meaning, and was further broken down into a ratio
for each individual and each tooth class.

```{r}
caries_count %>%
  group_by(class) %>% 
  caries_ratio(.caries = count)
```

As expected, the molars have a higher frequency of caries than the other teeth.

```{r}
caries_count %>% 
  group_by(id, class) %>% 
  caries_ratio(.caries = count)
```

```{r}
caries_count %>%
  caries_ratio(.caries = count, id, class)
```

```{r}
caries_count %>%
  caries_ratio(.caries = count, class, id) %>%
  ggplot(aes(x = class, y = ratio)) +
    geom_violin(aes(fill = class)) +
    geom_boxplot(width = 0.1)
```

```{r}
upper_order <- c(paste0("t", 18:11), paste0("t", 21:28))
lower_order <- c(paste0("t", 48:41), paste0("t", 31:38))

maxilla <- caries_count %>%
  filter(region == "maxilla") %>% 
  mutate(tooth = factor(tooth, levels = c(upper_order, lower_order))) %>% 
  group_by(tooth) %>% 
  summarise(
    n_teeth = n(),
    count = sum(count, na.rm = T),
    rate = count / n_teeth
    ) %>% 
  ggplot(aes(x = tooth, y = rate)) +
    geom_col(fill = "white") +
    theme_dark() +
    scale_y_reverse(limits = c(0.33,0), sec.axis = sec_axis(~.)) +
    scale_x_discrete(position = "top") +
    theme(
      axis.title.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.line = element_line(colour = "red", size = 1),
      axis.line.x.top = element_line(colour = "red", size = 4)
      ) +
  labs(y = "")

mandible <- caries_count %>%
  filter(region == "mandible") %>% 
  mutate(tooth = factor(tooth, levels = c(upper_order, lower_order))) %>% 
  group_by(tooth) %>% 
  summarise(
    n_teeth = n(),
    count = sum(count, na.rm = T),
    rate = count / n_teeth
    ) %>% 
  ggplot(aes(x = tooth, y = rate)) +
    geom_col(fill = "white") +
    scale_y_continuous(sec.axis = sec_axis(~.)) +
    theme_dark() +
    theme(axis.line = element_line(colour = "red", size = 1),
          axis.line.y.right = element_line(colour = "red", size = 1),
          axis.line.x.bottom = element_line(colour = "red", size = 6),
          axis.ticks.x = element_blank()) +
    labs(y = "Caries rate")

maxilla / mandible + plot_layout(guides = "collect")
```

### Periodontitis

```{r}
#| label: fig-periodont
dental_long %>%
  filter(!is.na(periodont)) %>% 
  dental_plot(fill = periodont)
```

### Calculus

```{r}
calc_index <- calculus_full %>%
  dental_longer(-id) %>% 
    #   group_by(id, quadrant) %>%
    # summarise(
    #   n_surf = sum(!is.na(score)),
    #   score_sum = sum(score, na.rm = T),
    #   calc_index = score_sum / n_surf
    # )

  calculus_index()
```

Distribution of calculus indices within the sample.
<!-- it is not calculating per surface (I think) -->

```{r}
#| label: fig-calculus-index
calc_index %>%
  ggplot(aes(x = quadrant, y = calc_index)) +
    geom_violin(aes(fill = quadrant), alpha = 0.6) +
    geom_boxplot(width = 0.1) +
    theme_minimal() +
    theme(panel.grid.major.x = element_blank())
```

```{r}
calc_index %>%
  left_join(uhplc_calculus_long, by = "id") %>%
  filter(
    batch == "batch1",
    compound != "thc",
    compound != "cbd",
    compound != "thcva",
    quadrant == "LA"
    ) %>% 
  ggplot(aes(x = as.factor(presence), y = calc_index)) +
    geom_violin(aes(fill = as.factor(presence)), alpha = 0.6) +
    geom_boxplot(width = 0.1) +
    facet_wrap(~ compound)
```

No apparent influence of lower anterior calculus index on the presence/absence
of a compound (or vice versa).

## Pathological conditions

Data on chronic maxillary sinusitis from @casnaUrbanizationRespiratory2021 were
included in this study to assess the relationship between upper respiratory
diseases with environmental factors (i.e. tobacco smoke, caffeine consumption).
Chronic maxillary sinusitis (CMS) is the inflammation of the lower paranasal
sinuses, air-filled pockets located in the skull that defend the organism against
inhaled particulate matter and pathogens. This occurs through the production of
mucus carried by small hairs toward an opening situated on the superior part of
the sinus, where pathogens are drained [@slavinDiagnosisManagement2005]. Without
drainage, mucus begins to accumulate in the sinuses, providing an ideal environment
for bacterial growth and thereby contributing to inflammation of the mucous
membranes and subsequently of the bone surfaces [@jangBoneInvolvement2002].
Lesions associated with CMS as defined by @boocockMaxillarySinusitis1995 were
recorded for each individual and classified as "pitting", "spicule-type bone
formation", "remodeled spicules", or "white pitted bone". CMS was scored as absent
when the sinus presented smooth surfaces with little or no associated pitting.
To facilitate inspection, fragmented sinuses were cleaned using a dry tooth-brush
and water where necessary. If the sinuses were not observable with the naked eye,
they were examined with a flexible medical endoscope (Pentax, model: FNL-10RBS,
ø=4mm; view angle=30°) inserted through minor breaks naturally occurring on the
inferior nasal conchae and palatine bone, where the bone tissue is thinner.

<!-- description of other diseases
Osteoarthritis
Vertebral osteophytosis
Cribra orbitalia
Degenerative disc disease
Schmorl's nodes
-->

## Statistical analysis

### Point-biserial correlation

Point-biserial (Pearson) correlation was conducted on compound concentrations,
calculus index, caries ratio, and binary variables.
This is done to see if any correlations exist prior to discretisation of continuous
variables.

```{r}
#| label: fig-pearson-corr
conc_cor %>%
  as_tibble(rownames = "var") %>% 
  mutate(across(!var, ~ if_else(.x == 1, NaN, .x))) %>%
  column_to_rownames("var") %>% 
ggcorrplot::ggcorrplot()
```

### Polychoric correlation

The calculus index and caries ratio for each individual was converted to an
ordinal variable by using quartiles, providing a score from 0--4.

```{r caries-calculus-quartiles}
caries_discrete
calculus_discrete
```

```{r polycorr}
polycorr$rho
```

```{r fig-polycorr}
ggcorrplot::ggcorrplot(polycorr$rho, type = "lower", show.diag = F)
```

The sample bias caused by targeting larger calculus deposits may also affect the detection of
compounds, as caffeinated drinks are often
acidic, and may cause a lower calculus formation.
We only found a weak negative correlation
between caffeine concentration and dental calculus index, but since we targeted individuals with
calculus, this may inadvertently have been controlled for during sample collection.

<!-- correlations on caries and calculus separated by quadrant/tooth type -->