### 2023-11-20 analysis/paper/paper.qmd

```{r}
#| label: setup
#| include: false
#| cache: false

library(here)
library(knitr)
#library(readr)
#devtools::load_all()
source(here("analysis/scripts/setup-qmd.R"))

options(ggplot.discrete.colour = function() scale_fill_viridis_d())
knitr::opts_chunk$set(
  fig.show = "hide",
  fig.path = "../figures/plos-",
  dev = "tiff",
  dev.args = list("tiff" = list(compression = "lzw"))
  #fig.keep = "none"
)
```

<!-- Keywords: {{< meta keywords >}} -->

<!-- Highlights: {{< meta highlights >}} -->

## Introduction

{{< include _intro.qmd >}}

## Materials

{{< include _materials.qmd >}}

## Methods

{{< include _methods.qmd >}}

## Results

{{< include _results.qmd >}}

## Discussion

{{< include _discussion.qmd >}}

## Conclusions

{{< include _conclusions.qmd >}}

## Acknowledgements {-}

We wish to thank Kirsten Ziesemer for helping track down the calculus samples
from her studies. Additionally, we owe special thanks to Vincent Falger and Kees
de Groot from the Middenbeemster Historical Society for their input on early
draft manuscripts and a lovely guided tour of Middenbeemster. We also thank
Louise Le Meillour for taking on our preprint and providing helpful comments,
and Mario Zimmerman and two other anonymous reviewers for their time and invaluable
insights.

## Funding {-}

This research has received funding from the European Research Council under the
European Union’s Horizon 2020 research and innovation program, grant agreement
number STG–677576 (“HARVEST”).

## Conflict of interest disclosure {-}

The authors have no conflicts to declare.

## Data Availability Statement {-}

All raw data is available on Zenodo (<https://doi.org/10.5281/zenodo.8061483>).
Analysis scripts, and the source code for the manuscript and
supplementary materials are available as a research compendium
(<https://doi.org/10.5281/zenodo.7649824>)
using the structure recommended by the **rrtools** R package [@Rrrtools].

## References {-}

::: {#refs}
:::