### 2024-04-24 analysis/paper/paper.qmd
incoporate changes from pcj branch

```{r}
#| label: setup
#| include: false
#| cache: false

library(here)
library(knitr)
#library(readr)
#devtools::load_all()
source(here("analysis/scripts/setup-qmd.R"))

options(ggplot.discrete.colour = function() scale_fill_viridis_d())
```

<!-- Keywords: {{< meta keywords >}} -->

<!-- Highlights: {{< meta highlights >}} -->

## Introduction

{{< include _intro.qmd >}}

## Materials

{{< include _materials.qmd >}}

## Methods

{{< include _methods.qmd >}}

## Results

{{< include _results.qmd >}}

## Discussion

{{< include _discussion.qmd >}}

## Conclusions

{{< include _conclusions.qmd >}}

## Acknowledgements {-}

We wish to thank Kirsten Ziesemer for helping track down the calculus samples
from her studies. Additionally, we owe special thanks to Vincent Falger and Kees
de Groot from the Middenbeemster Historical Society for their input on early
draft manuscripts and a lovely guided tour of Middenbeemster.

We also thank Louise Le Meillour for taking on our preprint and providing helpful
comments, and Mario Zimmerman and two other anonymous reviewers for their time
and invaluable insights. This paper was evaluated during a CODECHECK event at TU
Delft by Daniel Nüst and Stephen Eglen, who provided helpful feedback on the
reproducibility of all code used in this paper.

Preprint version 5 of this article has been peer-reviewed and recommended by PCI
Archaeology <https://doi.org/10.24072/pci.archaeo.100389> [@meillourDetectionPlantderived2023].

## Funding {-}

This research has received funding from the European Research Council under the
European Union’s Horizon 2020 research and innovation program, grant agreement
number STG–677576 (“HARVEST”).

## Conflict of interest disclosure {-}

The authors have no conflicts to declare.

## Data Availability Statement {-}

All raw data is available on Zenodo [<https://doi.org/10.5281/zenodo.8061483>,
@bartholdyDentalPathological2023].
Analysis scripts, and the source code for the manuscript and
supplementary materials are available as a research compendium
[<https://doi.org/10.5281/zenodo.7649824>, @bartholdyMB11CalculusPilot]
using the structure recommended by the **rrtools** R package [@Rrrtools].

## References {-}

::: {#refs}
:::