### 2025-02-13 README.Rmd
update README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)

Title <- "Multiproxy analysis exploring patterns of diet and disease in dental calculus and skeletal remains from a 19th century Dutch population"
```

# Research compendium for Bartholdy et al. 2023

[![R-CMD-check](https://github.com/bbartholdy/mb11CalculusPilot/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/bbartholdy/mb11CalculusPilot/actions/workflows/R-CMD-check.yaml) [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.7649825.svg)](https://zenodo.org/badge/DOI/10.5281/zenodo.7649825)

This repository contains the data and code for our paper:

> Bartholdy, Bjørn Peare; Hasselstrøm, Jørgen B.; Sørensen, Lambert K.; Casna, Maia; Hoogland, Menno; Historisch Genootschap Beemster; Henry, Amanda G. Multiproxy analysis exploring patterns of diet and disease in dental calculus and skeletal remains from a 19th century Dutch population. Peer Community Journal, Volume 4 (2024), article no. e46. doi : 10.24072/pcjournal.414. https://peercommunityjournal.org/articles/10.24072/pcjournal.414/

## Contents

The **root** directory contains:

  - [:file\_folder: data-raw](/data-raw/): R script to download and process the raw data from Zenodo (https://zenodo.org/record/8061483).
  - [:file\_folder: data](/data/): Data output from `data-raw/*.R`, used in the analysis (*.rda*).
  - [:file\_folder: R](/R/): R functions required in the analysis.

The **analysis** directory contains:

  - [:file\_folder: paper](/analysis/paper): Quarto source document
    for manuscript. Includes code to reproduce the figures and tables
    generated by the analysis.
  - [:file\_folder: figures](/analysis/figures): Plots and other
    illustrations
  - [:file\_folder:
    supplementary-materials](/analysis/supplementary-materials):
    Supplementary materials including notes and other documents
    prepared and collected during the analysis.

## How to reproduce the paper

**Software requirements:**

- [R](https://cran.r-project.org/)
- [Quarto](https://quarto.org/)
- [Git](https://git-scm.com/downloads) (optional)

**Linux dependencies:**

`libx11-dev libicu-dev make zlib1g-dev libcurl4-openssl-dev libssl-dev libfontconfig1-dev libfreetype6-dev libfribidi-dev libharfbuzz-dev libjpeg-dev libpng-dev libtiff-dev pandoc libxml2-dev libgit2-dev`

You can download the compendium as a zip from from this URL:

[main.zip](/archive/refs/heads/main.zip).

or clone the repository:

`git clone https://github.com/bbartholdy/mb11CalculusPilot.git`

Then:

- Open the folder in your preferred IDE
- run `renv::restore()` to install the required R packages
- run `quarto::quarto_render("analysis/paper/paper.qmd", output_format = "pdf")` in the R console (requires installing **quarto** R package) or `quarto render analysis/paper/paper.qmd --to pdf` to produce a PDF
- find the pdf in `./analysis/paper/_output`

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.